---
title: "Finite point-source diffusion model"
fontsize: 12pt
date: "03_04_19"
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)
library(knitr)
library(kableExtra)
library(patchwork)
library(VGAM)
library(nls.multstart)


knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../tools/echem_processing_tools.R")
source("../tools/plotting_tools.R")

theme_set(theme_1())
```

# A new diffusion model

Based on the explanation from this [Dartmouth link](http://www.dartmouth.edu/~cushman/courses/engs43/Chapter2.pdf) under "Initial punctual release between two boundaries" I got the following equation for diffusion between two barriers at $x = 0$ and $x = L$ for a point source at $x = a$.

$$ C(x,t) = \frac{M}{\sqrt{4 \pi D t}} 
\sum_{n = -\infty}^{\infty} \left[ \exp{ \left( \frac{-(x-2nL-a)^2}{4 D t} \right) } + 
\exp{ \left(  \frac{-(x-2nL+a)^2}{4 D t} \right) } \right] $$        

For our system, we will assume that the point source is at $a = 0$, so we can evaluate and simplify:

$$ C(x,t) = \frac{M}{\sqrt{4 \pi D t}} 
\sum_{n = -\infty}^{\infty} \left[ \exp{ \left( \frac{-(x-2nL)^2}{4 D t} \right) } + 
\exp{ \left(  \frac{-(x-2nL)^2}{4 D t} \right) } \right] $$   

$$ C(x,t) = \frac{M}{\sqrt{4 \pi D t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(x-2nL)^2}{4 D t} \right) } \right] $$   

This expression looks a little nasty, but it's pretty straightforward to write as a function. The function below will take in the parameters $M_0$, $L$, $D$, and calculate the concentration $C$ for a range of $x$ and $t$ values.


```{r}
finite_point_x_t <- function(m0, L, D, x, t){
  sum = 0
  for (n in seq(-100,100,1)) {
    sum = sum +  2*exp((-(x-2*n*L)^2)/(4*D*t))
  }
  (m0 / sqrt(4 * pi * D * t)) * sum 
}
```

So this function will calculate the concentration gradient over time! Here I'm calculating the first 200 terms, which should be sufficient.

# Generate some data

Now let's generate some a grid of time $t$ and space $x$, and calculate the concentration gradient for some arbitrary parameters $M_0 = 2$, $L = 0.1$, and $D = 5*10^{-5}$. 

```{r}
grid <- expand.grid(
  x = seq(0,0.1,length=101),
  t = seq(0, 2500, length = 25)
  )

finite_point_data <- grid %>% 
  mutate(c=finite_point_x_t(m0 = 2, L = 0.1, D = 5e-6, x=x,t=t ))
```

Ok, let's see what our concentration gradients look like!

```{r}
ggplot(finite_point_data, aes(x = x, y = c, color = t)) + geom_path(aes(group = t)) + scale_color_viridis()
```

Cool, so this shows that the concentration starts out high at $x = 0$ and slowly equilibrates to an almost flat gradient (at longer times it will be totally flat).

Ok, for our electrode system, we really only have a measurement at $x = 0$, so let's see what the concentration does over time at that point in space.
```{r}
ggplot(finite_point_data %>% filter(x==0), aes(x = t, y = c, color = t)) + geom_point() + scale_color_viridis()
```
Great! This looks similar to our data. This is equivalent to evaluating the expression above at $x = 0$, which gives:

$$ C(x,t) = \frac{M}{\sqrt{4 \pi D t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D t} \right) } \right] $$   

I'll write a function to do just that math and regenerate our data for just $x = 0$. Let's add a little noise too, to make it more realistic and also allow us to do a nonlinear fit.
```{r}
finite_point_x0 <- function(m0, L, D, t){
  sum = 0
  for (n in seq(-100,100,1)) {
    sum = sum +  2*exp((-(-2*n*L)^2)/(4*D*t))
  }
  (m0 / sqrt(4 * pi * D * t)) * sum
}

finite_point_x0_data <- grid %>% 
  filter(x ==0) %>% 
  mutate(c=finite_point_x0(m0 = 2, L = 0.1, D = 5e-6 ,t=t )) %>% 
  mutate(c_noise = rnorm(n=length((grid %>% filter(x==0))$t),mean=c,sd=1)) 
```

# Fitting for parameters

So here's what the concentration looks like at $x = 0$ over time with a little noise added:

```{r}

ggplot(finite_point_x0_data, aes(x = t, y = c_noise, color = t)) + geom_point() + scale_color_viridis()

```

Ok, now let's see if we can get the nonlinear least squares (nls) function to fit this data and give us back the parameters we used to generate the data.

```{r}


suppressWarnings(
  fit_nls_multstart <- nls_multstart(formula = c_noise~finite_point_x0(m0=2, L, D, t=t), 
            data = finite_point_x0_data,
            start_lower = c(L = 0.01, D = 1e-8),
            start_upper = c(L = 10, D = 1e-5),
            lower = c(L = 0, D = 0),
            supp_errors = 'Y',
            iter = 250,
            na.action = na.omit
            )
)


L_fit <- coef(fit_nls_multstart)[1]
D_fit <- coef(fit_nls_multstart)[2]

finite_point_x0_data_pred <- finite_point_x0_data %>% 
  mutate(pred = finite_point_x0(m0=2, L = L_fit, D = D_fit, t=t))

ggplot(finite_point_x0_data_pred %>% filter(t>0), aes(x = t, y = c_noise, color = t)) + 
  geom_line(aes(y = pred), color = 'black', linetype = 2) + 
  geom_point() + 
  scale_color_viridis()

summary(fit_nls_multstart)
```

Wow, it actually works! Note that we asked the nls to find the values of D and L, but we provided $M_0$. This is similar to the way I'm doing the analysis now, but it does depend on having an accurate guess for $M_0$. Let's see what estimate we get if $M_0$ is off by 2x.

```{r}

### m0 = 1 instead of m0 = 2

suppressWarnings(
  fit_nls_multstart <- nls_multstart(formula = c_noise~finite_point_x0(m0=1, L, D, t=t), 
            data = finite_point_x0_data,
            start_lower = c(L = 0.01, D = 1e-8),
            start_upper = c(L = 10, D = 1e-5),
            lower = c(L = 0, D = 0),
            supp_errors = 'Y',
            iter = 250,
            na.action = na.omit
            )
)


L_fit <- coef(fit_nls_multstart)[1]
D_fit <- coef(fit_nls_multstart)[2]

finite_point_x0_data_pred <- finite_point_x0_data %>% 
  mutate(pred = finite_point_x0(m0=1, L = L_fit, D = D_fit, t=t))

ggplot(finite_point_x0_data_pred %>% filter(t>0), aes(x = t, y = c_noise, color = t)) + 
  geom_line(aes(y = pred), color = 'black', linetype = 2) + 
  geom_point() + 
  scale_color_viridis()

summary(fit_nls_multstart)
```

So now the function still returns a nice looking best fit line, but the estimate of L is off 2x and the estimate of D is off almost 5x. 

I think that we can continue to deal with this in the same way as before, by estimating D with a high and low estimate of $I_0$ using the soak peak current and the first transfer SWV peak current. 

Now, I need to re adapt this math to the SWV current, and try to fit some of our real data!

