---
title: "Finite diffusion with SWV data"
fontsize: 12pt
date: "03_05_19"
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)
library(knitr)
library(kableExtra)
library(patchwork)
library(VGAM)
library(nls.multstart)


knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../tools/echem_processing_tools.R")
source("../tools/plotting_tools.R")

theme_set(theme_1())
```

# Math - finite diffusion with SWV data

So from the 03/04/19 notebook we have an expression for the concentration at $x = 0$ over time for a system with finite boundaries at $x = 0$ and $x = L$:

$$ C(x=0,t) = \frac{M}{\sqrt{4 \pi D t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D t} \right) } \right] $$  

We're going to adapt this slightly by normalizing by the cross sectional area to get an actual concentration in mol / cm^3 , and let's go ahead and label M0 and Dm. 

$$ C(t) = \frac{M_0}{A \sqrt{4 \pi D_m t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D_m t} \right) } \right] $$  

Recall that SWV current can be defined. 

$$ I_{swv} = \frac{n F A \sqrt{D_{ap}} C}{\sqrt{\pi t_p}} \psi $$
and so $C$ can be defined in terms of SWV parameters/data:

$$C = \frac{I_{swv} \sqrt{\pi t_p} }{n F A \sqrt{D_{ap}} \psi}$$

And so can $M_0$:

$$C = \frac{M_0}{V} = \frac{I_0 \sqrt{\pi t_p} }{n F A \sqrt{D_{ap}} \psi}$$

$$ M_0 = \frac{I_0 V \sqrt{\pi t_p} }{n F A \sqrt{D_{ap}} \psi}$$ 

Now we can substitute $C$ and $M_0$ with the equivalent SWV expressions:

$$ \frac{I_{swv} \sqrt{\pi t_p} }{n F A \sqrt{D_{ap}} \psi} = \frac{\frac{I_0 V \sqrt{\pi t_p} }{n F A \sqrt{D_{ap}} \psi}}{A \sqrt{4 \pi D_m t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D_m t} \right) } \right] $$  

$$ I_{swv}(t) = \frac{I_0 V}{A \sqrt{4 \pi D_m t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D_m t} \right) } \right] $$ 

Recall that $\frac{V}{A} = z$, which is the diffusion layer at the electrode.

$$ z = \sqrt{D_{ap} t_s}$$

So our expression is:

$$ I_{swv}(t) = \frac{I_0 \sqrt{D_{ap} t_s}}{ \sqrt{4 \pi D_m t}} 
\sum_{n = -\infty}^{\infty} \left[ 2 \exp{ \left( \frac{-(-2nL)^2}{4 D_m t} \right) } \right] $$ 

So we will need to provide our function to calculate $I_{swv}(t)$ with $D_{ap}$, $t_s$, and $I_0$. We will be fitting for unknowns $D_m$ and $L$.

# Fit some data
Before we even think about real data, let's right our function to calculate the decay curve from the electrochemical parameters using the above equation. 

```{r}
# Function to calculate decay curves from parameters - from eq. above. 

diffusion_point_finite <- function(t, i0, D_ap, t_s,  L, D_phys){
  sum = 0
  
  for (n in seq(-100,100,1)) {
    sum = sum +  exp((-(-2*n*L)^2)/(4*D_phys*t))
  }
  
  (2*i0 * sqrt(D_ap * t_s) / sqrt(4 * D_phys * t)) * sum
}


diffusion_point_semi <- function(t, i0, D_ap, t_s, D_phys){

  (i0 * sqrt(D_ap * t_s) / sqrt(4  * D_phys * t))
}
```

Now let's try to generate some data from reasonable parameters and see what it looks like. 

```{r}
df <- tibble(t = seq(0, 2500, length = 25)) %>% 
  mutate(sig1 = diffusion_point_finite(t=t, i0 = 1e-5,D_ap = 6e-6,t_s = 0.005 , L = 1, D_phys = 6e-8)) %>% 
  mutate(sig2 = diffusion_point_finite(t=t, i0 = 1e-5,D_ap = 6e-6,t_s = 0.005 , L = 0.015, D_phys = 6e-8)) %>% 
  mutate(sig3 = diffusion_point_semi(t=t, i0 = 1e-5,D_ap = 6e-6,t_s = 0.005 , D_phys = 6e-8))

ggplot(df, aes(x = t)) + 
  geom_point(aes(y = sig1), color = 'red', size = 3) + 
  geom_point(aes(y = sig2), color = 'purple', size = 1) + 
  geom_point(aes(y = sig3), color = 'blue') + 
  ylim(0,NA)
```

Finally, we can start with the real data. Let's try to fit one of the blank decay curves first. 
```{r}

blank_data <- read_csv("../11_28_18_blank_IDA/Processing/11_28_18_swv_decays_Processed.csv") %>% 
  filter(PHZadded!='10uM')

ggplot(blank_data , aes(x = norm_time, y = signal_from_transfer, color = PHZadded))+
  geom_point()+geom_line(aes(y = i0), linetype = 2, color = 'black') + 
  geom_hline(yintercept = 0)+
  facet_wrap(~PHZadded,ncol = 4) + ylim(0,NA)
```

Could shifting the curve up and down give different D estimates? Probably should check, because these curves don't reach 0 because of some background...

```{r}
remaining peak height
100uM = 2.648e-7
75uM = 2.199e-7
50uM = 1.958e-7
25uM = 1.548e-7

resids <- tibble(
  PHZadded = c('100uM','75uM','50uM','25uM'),
  resid = c(2.648e-7,2.199e-7,1.958e-7,1.548e-7))

last signal
100uM = 8.6817e-7

bg_subtract: -(min - peak)
- (8.6817 - 2.648)

blank_data %>% filter(PHZadded=='75uM' & rep==50)


```


```{r}
blank_data_bg <- left_join(blank_data, resids, by = c('PHZadded')) %>% 
  group_by(PHZadded) %>% 
  mutate(signal_bg = signal_from_transfer - (min(signal_from_transfer)-resid)) %>% 
  mutate(i0_bg = i0-(min(signal_from_transfer)-resid))



blank_data_nls <- blank_data_bg %>% 
  filter(norm_time<2) %>% 
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls_multstart(data=., 
          formula = signal_bg~diffusion_point_semi(
            t=norm_time, 
            i0 = max(i0_bg), 
            D_ap = 6.88e-6,
            t_s = 0.02 , 
            D_phys = D_phys ), 
          start_lower = c(D_phys = 1e-8),
          start_upper = c(D_phys = 1e-5),
          lower = c( D_phys = 0),
          supp_errors = 'Y',
          iter = 250,
          na.action = na.omit),
      conf.int=T)
    )


blank_data_preds <- left_join(blank_data_bg, blank_data_nls, by = c('PHZadded')) %>% 
  group_by(PHZadded) %>% 
  mutate(preds = diffusion_point_semi(
    t=norm_time, 
    i0 = max(i0_bg), 
    D_ap = 6.88e-6,
    t_s = 0.02 , 
    D_phys = estimate)) %>% 
    mutate(pred_high = diffusion_point_semi(
      t=norm_time, 
      i0 = max(i0_bg), 
      D_ap = 6.88e-6,
      t_s = 0.02 , 
      D_phys = conf.high))  %>% 
    mutate(pred_low = diffusion_point_semi(
      t=norm_time, 
      i0 = max(i0_bg), 
      D_ap = 6.88e-6,
      t_s = 0.02 ,
      D_phys = conf.low)) 

ggplot(blank_data_preds %>% filter(norm_time<5), aes(x = norm_time, y = signal_bg, color = PHZadded)) + 
  geom_point()+
  geom_line(aes(y = preds), linetype = 2, color = 'black') + 
  geom_line(aes(y = pred_high), linetype = 2, color = 'gray') + 
  geom_line(aes(y = pred_low), linetype = 2, color = 'gray') + 
  facet_wrap(~PHZadded, scales = 'free')+
  ylim(0,NA)
```


```{r}
blank_data_nls %>% arrange(term)
```

Let's start by fixing L at a high number, which should essentially be semi-infinite diffusion
```{r}
blank_data_bg <- blank_data %>%
  group_by(PHZadded) %>% 
  mutate(signal_bg = signal_from_transfer - min(signal_from_transfer)/2) %>% 
  mutate(i0_bg = i0-min(signal_from_transfer)/2)



blank_data_nls <- blank_data_bg %>% 
  filter(norm_time<2) %>% 
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls_multstart(data=., 
          formula = signal_bg~diffusion_point_finite(
            t=norm_time, 
            i0 = max(i0_bg), 
            D_ap = 6.88e-6,
            t_s = 0.02 , 
            L = 100, 
            D_phys = D_phys ), 
          start_lower = c(D_phys = 1e-8),
          start_upper = c(D_phys = 1e-5),
          lower = c( D_phys = 0),
          supp_errors = 'Y',
          iter = 250,
          na.action = na.omit),
      conf.int=T)
    )


blank_data_preds <- left_join(blank_data_bg, blank_data_nls, by = c('PHZadded')) %>% 
  group_by(PHZadded) %>% 
  mutate(preds = diffusion_point_finite(
    t=norm_time, 
    i0 = max(i0_bg), 
    D_ap = 6.88e-6,
    t_s = 0.02 , 
    D_phys = estimate,
    L=100)) %>% 
    mutate(pred_high = diffusion_point_finite(
      t=norm_time, 
      i0 = max(i0_bg), 
      D_ap = 6.88e-6,
      t_s = 0.02 , 
      D_phys = conf.high,
      L=100))  %>% 
    mutate(pred_low = diffusion_point_finite(
      t=norm_time, 
      i0 = max(i0_bg), 
      D_ap = 6.88e-6,
      t_s = 0.02 ,
      D_phys = conf.low,
      L=100)) 

ggplot(blank_data_preds %>% filter(norm_time<5), aes(x = norm_time, y = signal_bg, color = PHZadded)) + 
  geom_point()+
  geom_line(aes(y = preds), linetype = 2, color = 'black') + 
  geom_line(aes(y = pred_high), linetype = 2, color = 'gray') + 
  geom_line(aes(y = pred_low), linetype = 2, color = 'gray') + 
  facet_wrap(~PHZadded, scales = 'free')+
  ylim(0,NA)
```

Here I used bg subtracted data with the min / 2, because if you go back and look at the processing of the data for electrode i1, there's both background and signal left. Then I also changed t_s to give reasonable numbers. I think if we are going to use this method we should explicitly set Dap = Dphys and solve for t_s...but idk how to feel about this.

```{r}
blank_data_nls %>% arrange(term)
```

Let's see what happens if we let $L$ vary freely:

```{r}


blank_data_nls <- blank_data_bg %>% 
  filter(norm_time<3) %>% 
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls_multstart(data=., 
          formula = signal_from_transfer~diffusion_point_finite(
            t=norm_time, 
            i0 = max(i0), 
            D_ap = 6.88e-6,
            t_s = 0.02 , 
            L = L, 
            D_phys = D_phys ), 
          start_lower = c(L = 0.001, D_phys = 1e-8),
          start_upper = c(L=1, D_phys = 1e-5),
          lower = c(L = 0, D_phys = 0),
          supp_errors = 'Y',
          iter = 250,
          na.action = na.omit),
      )
    )




blank_data_preds <- left_join(blank_data_bg, blank_data_nls  %>% select(PHZadded,term,estimate) %>% spread(term, estimate,drop=T), by = c('PHZadded')) %>% 
  group_by(PHZadded) %>% 
  mutate(preds = diffusion_point_finite(
    t=norm_time, 
    i0 = max(i0), 
    D_ap = 6.88e-6,
    t_s = 0.02 , 
    L = L, 
    D_phys = D_phys)) 

ggplot(blank_data_preds %>% filter(norm_time<5), aes(x = norm_time, y = signal_from_transfer, color = PHZadded)) + 
  geom_point()+
  geom_line(aes(y = preds), linetype = 2, color = 'black') + facet_wrap(~PHZadded, scales = 'free')+
  ylim(0,NA)
```

```{r}
blank_data_nls %>% arrange(term)
```
Unfortunately this model will not fit the background subtracted data...unclear why.

We can add another term, $b$ to the equation, which can be interpreted as either background or perhaps a contribution from flow. This added complexity allows the model to fit extremely well, but it is unclear how realistic this equation is. 
```{r}
blank_data_bg <- blank_data %>%
  group_by(PHZadded) %>% 
  mutate(signal_bg = signal_from_transfer - min(signal_from_transfer))

blank_data_nls <- blank_data_bg %>% 
  filter(norm_time<3) %>% 
  group_by(PHZadded) %>% 
  do(
    tidy(
      nls_multstart(data=., 
          formula = signal_from_transfer~diffusion_point_finite(
            t=norm_time, 
            i0 = max(i0), 
            D_ap = 6.88e-6,
            t_s = 0.0053 , 
            L = L, 
            D_phys = D_phys )+b, 
          start_lower = c(L = 0.001, D_phys = 1e-8, b=-5e-7),
          start_upper = c(L=1, D_phys = 1e-5, b = 5e-7),
          lower = c(L = 0, D_phys = 0, b = -1),
          supp_errors = 'Y',
          iter = 250,
          na.action = na.omit),
      )
    )




blank_data_preds <- left_join(blank_data_bg, blank_data_nls  %>% select(PHZadded,term,estimate) %>% spread(term, estimate,drop=T), by = c('PHZadded')) %>% 
  group_by(PHZadded) %>% 
  mutate(preds = diffusion_point_finite(
    t=norm_time, 
    i0 = max(i0), 
    D_ap = 6.88e-6,
    t_s = 0.0053 , 
    L = L, 
    D_phys = D_phys) + b) 

ggplot(blank_data_preds %>% filter(norm_time<5), aes(x = norm_time, y = signal_from_transfer, color = PHZadded)) + 
  geom_point()+
  geom_line(aes(y = preds), linetype = 2) + facet_wrap(~PHZadded, scales = 'free')+
  ylim(0,NA)
```
You can see that the fits are now essentially perfect. 

The parameters we get are:
```{r}
blank_data_nls %>% arrange(term)
```
So this parameterization gives $D_{phys}$ in the 10^-7 cm^2 / sec range with a small L of ~10um and often a negative b. How unrealistic are these numbers? By allowing b to be negative are we artifically slowing D?

```{r}
ggplot(blank_data %>% filter(norm_time<3), aes(x = norm_time, y = signal_from_transfer- 8e-7, 
    color = PHZadded, shape = electrode)) + geom_point() + geom_smooth(method = "nls", 
    formula = y ~ b * (x)^-0.5 + a, method.args = list(start = c(b = 0.1, 
        a = 1e-07)), se = F) + facet_wrap(~electrode) + scale_color_viridis(discrete = T) + ylim(0,NA)
```