---
title: "Psoralen Toxicity "
subtitle: "Plate Reader Data"
fontsize: 12pt
date: "01_21_19 "
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../tools/echem_processing_tools.R")
source("../../tools/plotting_tools.R")

theme_set(theme_1())
```

# Purpose

The purpose of this experiment was to test the toxicity of trioxsalen toward *Pseudomonas aeruginosa* cells under growth conditions. I have been using trioxsalen to treat IDA biofilms, and I wanted to get a sense for whether toxicity would be a huge problem for that set of experiments. 

# Protocol 

deltaPHZstar or WT PA cells were grown in minimal succinate medium with 1uM or 10uM trioxsalen (from a 50x EtOH stock), or with just EtOH, or with nothing added at all. This was to separate the effect of the 0.5% Ethanol on the cultures. 

Each well contained the following: 

* 189 MM Succinate Medium
* 1uL EtOH / Psoralen or None
* 10uL $OD_{500}=0.1$ cells in MM
* 50uL sterile mineral oil on top

The plate was incubated for ~24hrs @ 37 degrees with medium shaking, reading OD 500 and fluorescence excited at 365nm every 15min.


# Data

The well data is saved separately from the metadata for each well. Here I join them:
```{r}
pr_data <- read_csv("../Data/plate_reader/01_21_19_psoralen_toxicity_growth_curve.csv") %>% 
  gather(key = well, value = abs, -time) %>% 
  mutate(time = as.numeric(time) / 60)

pr_metadata <- read_csv("../Data/plate_reader/01_21_19_psoralen_platereader_metadata.csv")

df_pr <- left_join(pr_data,pr_metadata, by='well') 
```

And now we can plot the data:

```{r}

ggplot(df_pr, aes(x = as.numeric(time) / 60, y = abs, color = factor(psoralen))) + 
  geom_point(size=0.5) + 
  facet_wrap(~strain) + 
  scale_color_viridis(discrete = T) +
  xlab("Hours") + 
  ylab("Abs @ 500nm")
```
It looks like WT and dPHZstar grew a little differently in this experiment, but it does not seem like the trioxsalen affected the growth curves. 

# Conclusion

It does not appear that trioxsalen is overly toxic to PA cells, and 10uM is an appropriate concentration to use in the IDA experiments. 