---
title: "Growth Curve Fitting"
author: "Scott Saunders"
date: "01_24_20"
output:
  html_document:
    code_folding: hide
    self_contained: no
    theme: cosmo
    toc: yes
fontsize: 12pt
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
#library(dplyr)
library(tidyverse)
library(cowplot)
library(viridis)
library(knitr)
library(kableExtra)
library(modelr)
library(broom)
library(ggforce)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../IDA/tools/plotting_tools.R")

theme_set(theme_1())

```

Read in some sample data to get a nice and messy dataset. 

```{r}
meta <- read_csv("data/plateReader_Conditions.csv")
df <- read_csv("data/plateReaderToxicity.csv") %>% gather(key = 'well',value = 'abs',-time) %>% mutate(time = time*24)

df_meta <- left_join(df, meta, by = 'well')

ggplot(df_meta, aes(x = time, y = abs, color = amount)) + geom_point()+facet_wrap(~molecule)
```

From [this paper](https://aem.asm.org/content/56/6/1875), it suggests that the modified Gompertz equation is the way to go. They desribe the three parameters as follows: 

>The three phases of the growth curve can be described by three parameters: the maximum specific growth rate, $\mu_m$ is defined as the tangent in the inflection point; the lag time, $\lambda$, is defined as the x-axis intercept of this tangent; and the asymptote $A$ is the maximal value reached.

$$ y = A*exp \left(-exp \left(\frac{\mu_m}{A}(\lambda - t) + 1 \right)\right)$$

Let's try our first model fit. Let's start with a nice classic growth curve from the untreated sample:

```{r}
curve_1 <- df_meta %>% filter(molecule == 'none' & replicate == 1 & strain == 'WT')

ggplot(curve_1, aes(x = time, y = abs)) + geom_point()
```

Even this nice growth curve has some messiness to it, but let's see how we can do.

```{r}
ggplot(curve_1, aes(x = time, y = abs)) + geom_point() + 
    geom_smooth(method='nls',
              formula=y~A*exp(-exp((mu/A)*(lam - x) + 1)),
              method.args=list(start=c(A = 1, mu = 0.1, lam = 2)),
              se=F)

ggplot(curve_1 %>% filter(time<10), aes(x = time, y = abs)) + geom_point() + 
    geom_smooth(method='nls',
              formula=y~A*exp(-exp((mu/A)*(lam - x) + 1)),
              method.args=list(start=c(A = 1, mu = 0.1, lam = 2)),
              se=F)
```

```{r}
nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1)), data = curve_1, start = c(A = 1, mu = 0.1, lam = 2))

nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1)), data = curve_1 %>% filter(time<10), start = c(A = 1, mu = 0.1, lam = 2))
```


```{r}

df_nest <- df_meta %>% 
  filter(molecule == 'none' & strain == 'WT') %>% 
  group_by(replicate) %>% 
  nest()

fit_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1))+c, start = c(A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
}

df_models <- df_nest %>% 
  mutate(models = map(data, fit_gompertz_mod)) 

df_preds <- df_models %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_model_ests <- df_models %>% 
  mutate(ests = map(models, tidy, conf.int = T)) %>% 
  unnest(ests)

ggplot(df_preds, aes(x = time)) + geom_point(aes(y = abs), shape =21) + geom_path(aes(y = pred), color = 'blue') + facet_wrap(~replicate)

ggplot(df_model_ests, aes(x = replicate, y = estimate)) + geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term,scales = 'free')
```
Adding a 'c' parameter here helped the model be able to fit curves that don't start at zero.


Now let's scale up to the whole thing and see how it does!

```{r}
df_all_nest <- df_meta %>%
  filter(amount <10000) %>% #AQDS and RB work too, etbr <= 500 works and MB <=500 and spermidine <=1000
  group_by(strain, molecule, amount, replicate) %>% 
  nest()

ggplot(df_meta %>% filter(molecule == 'SpmDn'), aes(x = time, y = abs, color = factor(amount))) + geom_point()

df_models_all <- df_all_nest %>% 
  mutate(models = map(data, fit_gompertz_mod))

df_models_all <- df_all_nest %>% 
  mutate(models = safely(map(data,fit_gompertz_mod))$result)



df_models_all[[1]]

df_preds_all <- df_models_all %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_model_ests_all <- df_models_all %>% 
  filter(molecule == 'none') %>% 
  mutate(ests = map(models, tidy, conf.int = T)) %>% 
  unnest(ests)

ggplot(df_preds_all %>% filter(molecule == 'none'), aes(x = time)) + geom_point(aes(y = absorbance), shape =21) + geom_path(aes(y = pred), color = 'blue') + facet_wrap(strain~replicate)

ggplot(df_model_ests_all %>% filter(molecule == 'none'), aes(x = replicate, y = estimate, color = strain)) + geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term,scales = 'free')
```
Moving toward handling models that cannot be fit:

```{r}
  
df_models_all_2 <- df_all_nest %>% 
  mutate(fit_result = map(data, safely(fit_gompertz_mod))) %>%
  ungroup() %>% 
  mutate(fit_errors = map(fit_result, 'error')) %>% 
  mutate(models = map(fit_result,'result')) %>% 
  mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))


df_preds_all_2 <- df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)


df_model_ests_all_2 <- df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(est_result = map(models, safely(tidy), conf.int = T)) %>% 
  mutate(est_errors = map(est_result, 'error')) %>% 
  mutate(ests = map(est_result,'result')) %>% 
  mutate(est_error_occured = map(.x = ests, .f = is.null), est_error_occured = unlist(est_error_occured)) %>% 
  mutate(ests = ifelse(est_error_occured == T, map(models, tidy, conf.int = F), ests)) %>% 
  unnest(ests)


df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(gla = map(models, glance)) %>% 
  unnest(gla)

```

# Starting fresh here to try and work out shiny code

```{r}
df_1 <- read_csv("../../Growth_curve_fitting/data/plate_reader_complex_data.csv")
df_2 <- read_csv("data/plateReader_Conditions.csv")

df <- df_1 %>% gather(key = 'well',value = 'absorbance',-time)
        
df_meta <- left_join(df, df_2, by = 'well') %>% filter(molecule == 'none')

# group and nest the data
df_nest <- df_meta %>%
  group_by(strain, replicate) %>% 
  nest()
    
df_nest <- df %>%
  group_by(well) %>% 
  nest()    
        
# The nls function used to fit the data
fit_gompertz_mod <- function(df){
  
  mod <- nls(formula=absorbance~A*exp(-exp((mu/A)*(lam - time) + 1))+c, 
             start = c(A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
        }
        
# fit data using the above function ...do this safely in case of errors
df_models <- df_nest %>% 
            mutate(fit_result = map(data, safely(fit_gompertz_mod))) %>%
            ungroup() %>% 
            mutate(fit_errors = map(fit_result, 'error')) %>% 
            mutate(models = map(fit_result,'result')) %>% 
            mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))

df_models$fit_errors[[1]]
        
df_preds <- df_models %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map2(data, models, add_predictions)) %>% 
            unnest(preds)

p <- ggplot(data = df_preds, aes_string(x=df_preds$time, y = df_preds$absorbance)) + 
  geom_point() + geom_path(aes(y = df_preds$pred), color = 'blue')+ facet_wrap(~well)
        
        print(p)

```

# Bi-Gompertz fitting??

```{r}
df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_simple_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_complex_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

ggplot(df_simple, aes(x = time, y = absorbance)) + geom_point()

fit_bi_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(
    
    formula=absorbance~( A_1*exp(-exp((mu_1/A_1)*(lam_1 - time) + 1)) ) + 
                       ( A_2*exp(-exp((mu_2/A_2)*(lam_2 - time) + 1)) ) + c,
    start = c(A_1 = 0.5, mu_1 = 0.2, lam_1 = 4, c= 0.05, 
              A_2 = 0.2, mu_2 = 0.3, lam_2 = 5),
    
    data = df)
  
  mod
}
                       
                       

fit_gompertz_mod <- function(df){
  
  mod <- nls(formula=absorbance~ B*(A*exp(-exp((mu/A)*(lam - time) + 1))+c), 
             start = c(B =1, A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
        }

df_nest <- df_simple %>%
  group_by(well) %>% 
  nest()    

df_models <- df_nest %>% 
            mutate(fit_result = map(data, safely(fit_bi_gompertz_mod))) %>%
            ungroup() %>% 
            mutate(fit_errors = map(fit_result, 'error')) %>% 
            mutate(models = map(fit_result,'result')) %>% 
            mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))


df_preds <- df_models %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map2(data, models, add_predictions)) %>% 
            unnest(preds)

ggplot(data = df_preds, aes_string(x=df_preds$time, y = df_preds$absorbance)) + 
  geom_point() + geom_path(aes(y = df_preds$pred), color = 'blue')+ facet_wrap(~well)

df_model_ests <- df_models %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(est_result = map(models, safely(tidy), conf.int = T)) %>% 
  mutate(est_errors = map(est_result, 'error')) %>% 
  mutate(ests = map(est_result,'result')) %>% 
  mutate(est_error_occured = map(.x = ests, .f = is.null), est_error_occured = unlist(est_error_occured)) %>% 
  mutate(ests = ifelse(est_error_occured == T, map(models, tidy, conf.int = F), ests)) %>% 
  unnest(ests)

ggplot(df_model_ests, aes(x = well, y = estimate)) + 
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term, scales = 'free')

factor(df_simple$well)


levels(factor(df_simple$well))
```


```{r}

fit_bi_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(
    
    formula=absorbance~( A_1*exp(-exp((mu_1/A_1)*(lam_1 - time) + 1)) ) + 
                       ( A_2*exp(-exp((mu_2/A_2)*(lam_2 - time) + 1)) ) + c,
    start = c(A_1 = 0.5, mu_1 = 0.2, lam_1 = 3, c= 0.1, 
              A_2 = 0.2, mu_2 = 0.1, lam_2 = 6),
    
    data = df)
  
  mod
}

gen_gompertz <- function(a, mu, lam, c, time){
  
  a*exp(-exp((mu/a)*(lam - time) + 1 )) + c
  
}

df_pred_ests <- left_join(df_preds,
          df_model_ests %>% pivot_wider(id_cols = 'well', names_from = 'term', values_from = 'estimate' ), 
          by = 'well') %>% 
  mutate(pred_1 = gen_gompertz(a = A_1, mu = mu_1, lam = lam_1, c = c, time = time)) %>% 
  mutate(pred_2 = gen_gompertz(a = A_2, mu = mu_2, lam = lam_2, c = c, time = time)) 


p <- ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free')

ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free', page = 12)

pages <- n_pages(p)

for(i in pages){
  p <- ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free', page = i)
  
  print(p)
}

```
# Test out GroFit

```{r}

library(grofit)

df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_simple_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

dataset <- ran.data(d=30, t=50, mu=1, lambda = 5, A = 15)

data <- dataset$data

time <- dataset$time

test_run <- grofit(time, data, FALSE, control = grofit.control(interactive = F))

plot(test_run$gcFit)

plot(test_run$drBootSpline, xlim = c(0, 50), ylim = c(0,15))

test_spline <- gcFitSpline(1:50, time[1,], control = grofit.control(interactive = F))

df <- tibble(time = test_spline$raw.time, preds = test_spline$fit.data, rawY = test_spline$raw.data)

ggplot(data = df, aes(x = time, y = preds)) + geom_point() + geom_point(aes(y = rawY), color = 'red')

test_Bootspline <- gcBootSpline(1:50, time[1,], control = grofit.control(interactive = F, nboot.gc = 100))

df <- tibble(time = test_Bootspline$boot.time, preds = test_spline$fit.data, rawY = test_spline$raw.data)

ggplot(data = df, aes(x = time, y = preds)) + geom_point() + geom_point(aes(y = rawY), color = 'red')

plot(test_Bootspline)

x <- 1:200
y <- 1/(1+exp(0.5*(15-x)))+rnorm(30)/20
TestRun <- gcBootSpline(x,y,"ID",grofit.control(nboot.gc=50))
print(summary(TestRun))
plot(TestRun)

TestRun$raw.time

TestRun$raw.data

TestRun$boot.data[1,]

df_boot <- as.tibble(t(as.matrix(TestRun$boot.data))) %>% mutate(time = 1:25) %>% pivot_longer(cols = -time, names_to = 'boot_num', values_to = 'preds')

ggplot(mapping = aes(x = x, y = y)) + geom_point()

df <- tibble(time = TestRun$raw.time, data = TestRun$raw.data, boot1 = TestRun$boot.data[1,])

ggplot(df, aes(x = time, y = data)) + geom_point() + geom_path(aes(y = boot1))

df_boot <- as.tibble(t(as.matrix(TestRun$boot.data))) %>% mutate(time = 1:25, data = TestRun$raw.data) %>% pivot_longer(cols = -c(time, data), names_to = 'boot_num', values_to = 'preds')

ggplot(df_boot, aes(x = time, y = preds, group = boot_num)) + geom_path(alpha = 0.5) + geom_point(aes(y = data), color = 'red')

df <- gcFitSpline(time = x, data = y, 'test', grofit.control(interactive = F))


df_plot <- tibble(time = df$raw.time, data = df$raw.data, pred = df$fit.data )

ggplot(df_plot, aes(x = time, y = pred)) + geom_path() + geom_point(aes(y = data), color = 'red')


```

A working example with spline bootstrapping

```{r}
library(grofit)
library(tidybayes)
library(ggridges)

x <- 1:200
y <- 1/(1+exp(0.5*(15-x)))+rnorm(30)/20

df_boot <- gcBootSpline(time = x, data = y, 'test_boot', grofit.control(interactive = F, nboot.gc = 100))

df_boot_times <- as.tibble(t(df_boot$boot.time)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'time')

df_boot_preds <- as.tibble(t(df_boot$boot.data)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'pred')

df_boot_plot <- left_join(df_boot_times, df_boot_preds, by = c('index','boot_num'))

#full_join(as_tibble(df_boot$boot.time), as_tibble(df_boot$boot.data))

#as.tibble(t(df_boot$boot.data))

#df_boot_plot <- tibble(time = df_boot$raw.time, data = df_boot$raw.data) %>%  
 # bind_cols(., as.tibble(t(df_boot$boot.data))) %>% 
 # pivot_longer(cols = -c(time, data), names_to = 'boot_num', values_to = 'pred')

ggplot(df_boot_plot, aes(x = time, y = pred, group = boot_num)) + geom_path(alpha = 0.01)+ geom_point(data = tibble(x=x, y = y), aes(x,y, group = NULL), color = 'red')



ggplot(df_boot_plot, aes(x = time, y = pred)) + stat_lineribbon()

```

and now for the parameter estimation

```{r}
df_boot_params <- bind_cols(lambda = df_boot$lambda, mu = df_boot$mu, A = df_boot$A) %>% 
  pivot_longer(cols = c(lambda, mu, A),names_to = 'term',values_to = 'estimate') %>% 
  filter(!is.na(estimate))

ggplot(df_boot_params, aes(x = estimate, y = term)) +  geom_point()

ggplot(df_boot_params, aes(x = estimate, y = term)) + geom_halfeyeh() 


df = tribble(
  ~group, ~subgroup, ~value,
  "a",          "h", rnorm(500, mean = 5),
  "b",          "h", rnorm(500, mean = 7, sd = 1.5),
  "c",          "h", rnorm(500, mean = 8),
  "c",          "i", rnorm(500, mean = 9),
  "c",          "j", rnorm(500, mean = 7)
) %>%
  unnest(value)

# here are vertical eyes:
df %>%
  ggplot(aes(x = group, y = value)) +
  stat_eye()

df_boot_params %>% 
  ggplot(aes(x = term, y = estimate)) + stat_density() + geom_point(alpha = 0.1)+ geom_boxplot() + facet_wrap(~term, scales = 'free')

ggplot(df_boot_params, aes(x = estimate, y = term)) + stat_halfeyeh() + facet_wrap(~term, scales = 'free')

tibble(a = c(seq(1:9), NA), group = c('A','A','A','B','B','B','C','C','C','C')) %>% 
  ggplot() + stat_halfeyeh(aes(x = a, y = group))

ests <- df_boot_params %>% 
  group_by(term) %>% 
  summarise(median = median(estimate))

ggplot(df_boot_plot, aes(x = time, y = pred, group = boot_num)) + geom_path(alpha = 0.01)+ geom_point(data = tibble(x=x, y = y), aes(x,y, group = NULL), color = 'red') + 
  geom_abline(slope = ests$median[3], intercept = -ests$median[2] * ests$median[3]) + 
  geom_hline(yintercept = ests$median[1])

```

Try on one real growth curve.

```{r}
df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_simple_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

df_g1 <- df_simple %>% filter(well == 'G1-WT')

df_g1_boot <- gcBootSpline(time = df_g1$time, data = df_g1$absorbance, 'test_boot', grofit.control(interactive = F, nboot.gc = 100))


df_g1_boot_times <- as.tibble(t(df_g1_boot$boot.time)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'time')

df_g1_boot_preds <- as.tibble(t(df_g1_boot$boot.data)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'pred')

df_g1_boot_plot <- left_join(df_g1_boot_times, df_g1_boot_preds, by = c('index','boot_num'))

ggplot(df_g1_boot_plot, aes(x = time, y = pred, group = boot_num))  + geom_point(data = df_g1, aes(x = time,y = absorbance, group = NULL), color = 'red', shape = 21)+ geom_path(alpha = 0.01)


```
and the parameter estimates:

```{r}

df_g1_boot_params <- bind_cols(lambda = df_g1_boot$lambda, mu = df_g1_boot$mu, A = df_g1_boot$A) %>% 
  pivot_longer(cols = c(lambda, mu, A),names_to = 'term',values_to = 'estimate') %>% 
  filter(!is.na(estimate))

ggplot(df_g1_boot_params, aes(x = estimate, y = term)) + stat_halfeyeh() + facet_wrap(~term, scales = 'free')


ests <- df_g1_boot_params %>% 
  group_by(term) %>% 
  summarise(median = median(estimate), mean = mean(estimate))

ggplot(df_g1_boot_plot, aes(x = time, y = pred, group = boot_num)) + geom_path(alpha = 0.01)+ 
  geom_point(data = df_g1, aes(x = time,y = absorbance, group = NULL), color = 'red', shape = 21)+
  geom_abline(slope = ests$mean[3], intercept = -ests$mean[2] * ests$mean[3]) + 
  geom_hline(yintercept = ests$median[1])
```


Try on many real growth curves
```{r}
#df_boot <- gcBootSpline(time = x, data = y, 'test_boot', grofit.control(interactive = F, nboot.gc = 100))

df_nest <- df_simple %>%
  group_by(well) %>% 
  nest()    

#df_g1_boot <- gcBootSpline(time = df_g1$time, data = df_g1$absorbance, 'test_boot', grofit.control(interactive = F, nboot.gc = 100))

#df_g1_boot <- gcFitSpline(time = df_g1$time, data = df_g1$absorbance, 'test best', grofit.control(interactive = F))



fit_best_spline <- function(df){
  
  mod <- gcFitSpline(time = df$time, data = df$absorbance, gcID = 'best spline', control = grofit.control(interactive = F))
  
  mod
}

fit_boot_splines <- function(df){
  
  mod <- gcBootSpline(time = df$time, data = df$absorbance, gcID = 'bootstrapped splines', control = grofit.control(interactive = F, nboot.gc = 500))
  
  mod
}

df_models <- df_nest %>% 
  mutate(bestSpline = map(data, fit_best_spline)) %>% 
  mutate(bootSpline = map(data, fit_boot_splines))

tidy_best_spline_preds <- function(df){
  tibble(time = df$fit.time,
         pred = df$fit.data,
         absorbance = df$raw.data)
}

tidy_boot_spline_preds <- function(df){
  
  df_boot_times <- as.tibble(t(df$boot.time)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'time')
  df_boot_preds <- as.tibble(t(df$boot.data)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'pred')
  
  df_boot_plot <- left_join(df_boot_times, df_boot_preds, by = c('index','boot_num') )

  df_boot_plot
}

#df_g1_boot_times <- as.tibble(t(df_g1_boot$boot.time)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'time')

#df_g1_boot_preds <- as.tibble(t(df_g1_boot$boot.data)) %>% mutate(index = row_number()) %>% pivot_longer(cols = -index, names_to = 'boot_num',values_to= 'pred')

#df_g1_boot_plot <- left_join(df_g1_boot_times, df_g1_boot_preds, by = c('index','boot_num'))

# this step takes forever, for 6 growth curves and only 100 bootstraps the data frame is >100,000 observations 
# plot the spline model from the single gcFitSpline(), but get the parameter estimates from the bootstraps!
df_preds <- df_models %>% 
            mutate(preds = map(bestSpline, tidy_best_spline_preds)) %>% 
            unnest(preds)

head(df_preds)

# Lineribbon is not very informative because splines are very certain...
ggplot(data = df_preds, aes(x=time, y = pred)) + 
  geom_path() + facet_wrap(~well, scales = 'free') + 
  geom_point(aes(y = absorbance), shape = 21, color = 'red', alpha =0.1)

### Move on to extracting parameter estimates

tidy_boot_spline_ests <- function(df){
  
  tibble(Lag = df$lambda, Mu = df$mu, A = df$A) %>% 
  pivot_longer(cols = c(Lag, Mu, A),names_to = 'term',values_to = 'estimate') %>% 
  filter(!is.na(estimate))
  
}

tidy_best_spline_ests <- function(df){
  
  tibble(Lag = df$parameters$lambda, Mu = df$parameters$mu, A = df$parameters$A) %>% 
  pivot_longer(cols = c(Lag, Mu, A),names_to = 'term',values_to = 'estimate') %>% 
  filter(!is.na(estimate))
  
}

#df_g1_boot_params <- bind_cols(lambda = df_g1_boot$lambda, mu = df_g1_boot$mu, A = df_g1_boot$A) %>% 
#  pivot_longer(cols = c(lambda, mu, A),names_to = 'term',values_to = 'estimate') %>% 
#  filter(!is.na(estimate))

df_model_ests <- df_models %>% 
 # mutate(ests = map(bootSpline, tidy_boot_spline_ests )) %>% 
  mutate(ests = map(bestSpline, tidy_best_spline_ests )) %>% 
  unnest(ests)


head(df_model_ests)

# Plot parameter estimates - G3 is weird
ggplot(df_model_ests %>% filter(well != 'G3-WT'), aes(x = estimate, y = well)) + stat_halfeyeh() + facet_wrap(~term, scales = 'free')


df_model_ests_summary <- df_model_ests %>% 
  group_by(well,term) %>% 
  summarise(median = median(estimate)) %>% 
  pivot_wider(id_cols = well, names_from = term, values_from = median)

ggplot(data = df_preds, aes(x=time, y = pred)) + 
  geom_path() + facet_wrap(~well, scales = 'free') + 
  geom_point(aes(y = absorbance), shape = 21, color = 'red', alpha =0.1) + 
  geom_abline(data = df_model_ests_summary, aes(slope = Mu, intercept = -Lag * Mu)) + 
  geom_hline(data = df_model_ests_summary, aes(yintercept = A))
```










```{r}
sessionInfo()
```