---
title: "Growth Curve Fitting"
author: "Scott Saunders"
date: "01_24_20"
output:
  html_document:
    code_folding: hide
    self_contained: no
    theme: cosmo
    toc: yes
fontsize: 12pt
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(viridis)
library(knitr)
library(kableExtra)
library(modelr)
library(broom)
library(ggforce)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../IDA/tools/plotting_tools.R")

theme_set(theme_1())

```

Read in some sample data to get a nice and messy dataset. 

```{r}
meta <- read_csv("data/plateReader_Conditions.csv")
df <- read_csv("data/plateReaderToxicity.csv") %>% gather(key = 'well',value = 'abs',-time) %>% mutate(time = time*24)

df_meta <- left_join(df, meta, by = 'well')

ggplot(df_meta, aes(x = time, y = abs, color = amount)) + geom_point()+facet_wrap(~molecule)
```

From [this paper](https://aem.asm.org/content/56/6/1875), it suggests that the modified Gompertz equation is the way to go. They desribe the three parameters as follows: 

>The three phases of the growth curve can be described by three parameters: the maximum specific growth rate, $\mu_m$ is defined as the tangent in the inflection point; the lag time, $\lambda$, is defined as the x-axis intercept of this tangent; and the asymptote $A$ is the maximal value reached.

$$ y = A*exp \left(-exp \left(\frac{\mu_m}{A}(\lambda - t) + 1 \right)\right)$$

Let's try our first model fit. Let's start with a nice classic growth curve from the untreated sample:

```{r}
curve_1 <- df_meta %>% filter(molecule == 'none' & replicate == 1 & strain == 'WT')

ggplot(curve_1, aes(x = time, y = abs)) + geom_point()
```

Even this nice growth curve has some messiness to it, but let's see how we can do.

```{r}
ggplot(curve_1, aes(x = time, y = abs)) + geom_point() + 
    geom_smooth(method='nls',
              formula=y~A*exp(-exp((mu/A)*(lam - x) + 1)),
              method.args=list(start=c(A = 1, mu = 0.1, lam = 2)),
              se=F)

ggplot(curve_1 %>% filter(time<10), aes(x = time, y = abs)) + geom_point() + 
    geom_smooth(method='nls',
              formula=y~A*exp(-exp((mu/A)*(lam - x) + 1)),
              method.args=list(start=c(A = 1, mu = 0.1, lam = 2)),
              se=F)
```

```{r}
nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1)), data = curve_1, start = c(A = 1, mu = 0.1, lam = 2))

nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1)), data = curve_1 %>% filter(time<10), start = c(A = 1, mu = 0.1, lam = 2))
```


```{r}

df_nest <- df_meta %>% 
  filter(molecule == 'none' & strain == 'WT') %>% 
  group_by(replicate) %>% 
  nest()

fit_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(formula=abs~A*exp(-exp((mu/A)*(lam - time) + 1))+c, start = c(A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
}

df_models <- df_nest %>% 
  mutate(models = map(data, fit_gompertz_mod)) 

df_preds <- df_models %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_model_ests <- df_models %>% 
  mutate(ests = map(models, tidy, conf.int = T)) %>% 
  unnest(ests)

ggplot(df_preds, aes(x = time)) + geom_point(aes(y = abs), shape =21) + geom_path(aes(y = pred), color = 'blue') + facet_wrap(~replicate)

ggplot(df_model_ests, aes(x = replicate, y = estimate)) + geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term,scales = 'free')
```
Adding a 'c' parameter here helped the model be able to fit curves that don't start at zero.


Now let's scale up to the whole thing and see how it does!

```{r}
df_all_nest <- df_meta %>%
  filter(amount <10000) %>% #AQDS and RB work too, etbr <= 500 works and MB <=500 and spermidine <=1000
  group_by(strain, molecule, amount, replicate) %>% 
  nest()

ggplot(df_meta %>% filter(molecule == 'SpmDn'), aes(x = time, y = abs, color = factor(amount))) + geom_point()

df_models_all <- df_all_nest %>% 
  mutate(models = map(data, fit_gompertz_mod))

df_models_all <- df_all_nest %>% 
  mutate(models = safely(map(data,fit_gompertz_mod))$result)



df_models_all[[1]]

df_preds_all <- df_models_all %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)

df_model_ests_all <- df_models_all %>% 
  filter(molecule == 'none') %>% 
  mutate(ests = map(models, tidy, conf.int = T)) %>% 
  unnest(ests)

ggplot(df_preds_all %>% filter(molecule == 'none'), aes(x = time)) + geom_point(aes(y = absorbance), shape =21) + geom_path(aes(y = pred), color = 'blue') + facet_wrap(strain~replicate)

ggplot(df_model_ests_all %>% filter(molecule == 'none'), aes(x = replicate, y = estimate, color = strain)) + geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term,scales = 'free')
```
Moving toward handling models that cannot be fit:

```{r}
  
df_models_all_2 <- df_all_nest %>% 
  mutate(fit_result = map(data, safely(fit_gompertz_mod))) %>%
  ungroup() %>% 
  mutate(fit_errors = map(fit_result, 'error')) %>% 
  mutate(models = map(fit_result,'result')) %>% 
  mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))


df_preds_all_2 <- df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(preds = map2(data, models, add_predictions)) %>% 
  unnest(preds)


df_model_ests_all_2 <- df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(est_result = map(models, safely(tidy), conf.int = T)) %>% 
  mutate(est_errors = map(est_result, 'error')) %>% 
  mutate(ests = map(est_result,'result')) %>% 
  mutate(est_error_occured = map(.x = ests, .f = is.null), est_error_occured = unlist(est_error_occured)) %>% 
  mutate(ests = ifelse(est_error_occured == T, map(models, tidy, conf.int = F), ests)) %>% 
  unnest(ests)


df_models_all_2 %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(gla = map(models, glance)) %>% 
  unnest(gla)

```

# Starting fresh here to try and work out shiny code

```{r}
df_1 <- read_csv("../../Growth_curve_fitting/data/plate_reader_complex_data.csv")
df_2 <- read_csv("data/plateReader_Conditions.csv")

df <- df_1 %>% gather(key = 'well',value = 'absorbance',-time)
        
df_meta <- left_join(df, df_2, by = 'well') %>% filter(molecule == 'none')

# group and nest the data
df_nest <- df_meta %>%
  group_by(strain, replicate) %>% 
  nest()
    
df_nest <- df %>%
  group_by(well) %>% 
  nest()    
        
# The nls function used to fit the data
fit_gompertz_mod <- function(df){
  
  mod <- nls(formula=absorbance~A*exp(-exp((mu/A)*(lam - time) + 1))+c, 
             start = c(A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
        }
        
# fit data using the above function ...do this safely in case of errors
df_models <- df_nest %>% 
            mutate(fit_result = map(data, safely(fit_gompertz_mod))) %>%
            ungroup() %>% 
            mutate(fit_errors = map(fit_result, 'error')) %>% 
            mutate(models = map(fit_result,'result')) %>% 
            mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))

df_models$fit_errors[[1]]
        
df_preds <- df_models %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map2(data, models, add_predictions)) %>% 
            unnest(preds)

p <- ggplot(data = df_preds, aes_string(x=df_preds$time, y = df_preds$absorbance)) + 
  geom_point() + geom_path(aes(y = df_preds$pred), color = 'blue')+ facet_wrap(~well)
        
        print(p)

```

# Bi-Gompertz fitting??

```{r}
df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_simple_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

df_simple <- read_csv("../../Growth_curve_fitting/data/plate_reader_complex_data.csv") %>% 
  gather(key = 'well',value = 'absorbance',-time)

ggplot(df_simple, aes(x = time, y = absorbance)) + geom_point()

fit_bi_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(
    
    formula=absorbance~( A_1*exp(-exp((mu_1/A_1)*(lam_1 - time) + 1)) ) + 
                       ( A_2*exp(-exp((mu_2/A_2)*(lam_2 - time) + 1)) ) + c,
    start = c(A_1 = 0.5, mu_1 = 0.2, lam_1 = 4, c= 0.05, 
              A_2 = 0.2, mu_2 = 0.3, lam_2 = 5),
    
    data = df)
  
  mod
}
                       
                       

fit_gompertz_mod <- function(df){
  
  mod <- nls(formula=absorbance~ B*(A*exp(-exp((mu/A)*(lam - time) + 1))+c), 
             start = c(B =1, A = 1, mu = 0.2, lam = 3, c = 0.1), data = df)
  
  mod
        }

df_nest <- df_simple %>%
  group_by(well) %>% 
  nest()    

df_models <- df_nest %>% 
            mutate(fit_result = map(data, safely(fit_bi_gompertz_mod))) %>%
            ungroup() %>% 
            mutate(fit_errors = map(fit_result, 'error')) %>% 
            mutate(models = map(fit_result,'result')) %>% 
            mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))


df_preds <- df_models %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map2(data, models, add_predictions)) %>% 
            unnest(preds)

ggplot(data = df_preds, aes_string(x=df_preds$time, y = df_preds$absorbance)) + 
  geom_point() + geom_path(aes(y = df_preds$pred), color = 'blue')+ facet_wrap(~well)

df_model_ests <- df_models %>% 
  filter(fit_error_occured == FALSE) %>% 
  mutate(est_result = map(models, safely(tidy), conf.int = T)) %>% 
  mutate(est_errors = map(est_result, 'error')) %>% 
  mutate(ests = map(est_result,'result')) %>% 
  mutate(est_error_occured = map(.x = ests, .f = is.null), est_error_occured = unlist(est_error_occured)) %>% 
  mutate(ests = ifelse(est_error_occured == T, map(models, tidy, conf.int = F), ests)) %>% 
  unnest(ests)

ggplot(df_model_ests, aes(x = well, y = estimate)) + 
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + facet_wrap(~term, scales = 'free')

```


```{r}

fit_bi_gompertz_mod <- function(df){
  
  #A = 1; mu = 0.1; lam = 2
  
  mod <- nls(
    
    formula=absorbance~( A_1*exp(-exp((mu_1/A_1)*(lam_1 - time) + 1)) ) + 
                       ( A_2*exp(-exp((mu_2/A_2)*(lam_2 - time) + 1)) ) + c,
    start = c(A_1 = 0.5, mu_1 = 0.2, lam_1 = 3, c= 0.1, 
              A_2 = 0.2, mu_2 = 0.1, lam_2 = 6),
    
    data = df)
  
  mod
}

gen_gompertz <- function(a, mu, lam, c, time){
  
  a*exp(-exp((mu/a)*(lam - time) + 1 )) + c
  
}

df_pred_ests <- left_join(df_preds,
          df_model_ests %>% pivot_wider(id_cols = 'well', names_from = 'term', values_from = 'estimate' ), 
          by = 'well') %>% 
  mutate(pred_1 = gen_gompertz(a = A_1, mu = mu_1, lam = lam_1, c = c, time = time)) %>% 
  mutate(pred_2 = gen_gompertz(a = A_2, mu = mu_2, lam = lam_2, c = c, time = time)) 


p <- ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free')

ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free', page = 12)

pages <- n_pages(p)

for(i in pages){
  p <- ggplot(df_pred_ests, aes(x = time)) + 
    geom_point(aes(y = absorbance), shape = 21)+
  geom_path(aes(y = pred), color = 'blue', size = 1) + 
  geom_path(aes(y = pred_1), color = 'red', size = 1) + 
  geom_path(aes(y = pred_2), color = 'red', linetype = 2, size = 1) + 
  facet_wrap_paginate(~well, nrow = 2, ncol = 2, scales = 'free', page = i)
  
  print(p)
}

```


-------

```{r}
sessionInfo()
```