---
title: "Reduced Phenazines with ctDNA titration"
author: "Scott Saunders"
fontsize: 12pt
date: '10_09_19'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: hide
    self_contained: no
    toc: yes
  github_document:
    pandoc_args: --webtex
---

```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(viridis)
library(knitr)
library(kableExtra)
library(modelr)
library(broom)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../IDA/tools/plotting_tools.R")

theme_set(theme_1())
```

# Intro

Do reduced phenazines bind DNA in vitro? We would expect that their endogenous fluorescence would change upon binding. Will we see a change in fluorescence upon exposure to DNA?

# Methods

90uM of each phenazine, titrating in calf thymus DNA from 2mg/mL stock.

135uL reduced phenazine
15uL of PBS + ctDNA (two fold dilutions)

6 concentrations from 0 to 280uM ctDNA (final)

triplicate of each measurement

# Results

## Import data

```{r}
df_360 <- read_csv("data/2019_10_09_PHZred_top_360-460.csv") %>% 
  gather(key = 'column', value = 'intensity', -row)

df_485 <- read_csv("data/2019_10_09_PHZred_top_485-528.csv") %>% 
  gather(key = 'column', value = 'intensity', -row)

df_meta <- read_csv("data/2019_10_09_PHZred_well_metadata.csv") %>% 
  gather(key = 'column', value = 'id', -row)

df_360_meta <- left_join(df_360, df_meta, by = c('row','column')) %>% 
  mutate(dna = case_when(
    column %in% c(1,7) ~ 0,
    column %in% c(2,8) ~ 17.5,
    column %in% c(3,9) ~ 35,
    column %in% c(4,10) ~ 70,
    column %in% c(5,11) ~ 140,
    column %in% c(6,12) ~ 280,
  ))

df_485_meta <- left_join(df_485, df_meta, by = c('row','column')) %>% 
    mutate(dna = case_when(
    column %in% c(1,7) ~ 0,
    column %in% c(2,8) ~ 17.5,
    column %in% c(3,9) ~ 35,
    column %in% c(4,10) ~ 70,
    column %in% c(5,11) ~ 140,
    column %in% c(6,12) ~ 280,
  ))
```


## 360nm ex

```{r}
ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity)) + geom_point() + facet_wrap(~id, scales = 'free')

ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity, color = id)) + geom_point() + geom_smooth(method = 'lm')
```

```{r}
#ggplot(df_360_meta %>% filter(id=='empty'), aes(x = as.integer(column), y = intensity, color = row)) + geom_point()

#ggplot(df_360_meta %>% filter(id=='PBS'), aes(x = as.integer(column), y = intensity, color = row)) + geom_point()

#ggplot(df_360_meta %>% filter(id %in% c('empty','PBS')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()
```

```{r}
ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PCA')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

ggplot(df_360_meta %>% filter(id %in% c('empty','PBS','PCN')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

ggplot(df_360_meta %>% filter(id %in% c('control','PBS','PYO')), aes(x = as.integer(column), y = intensity, color = id)) + geom_point()

```

```{r}
#ggplot(df_360_meta %>% filter(id %in% c('PYO')), aes(x = dna, y = intensity, color = id)) + geom_point()

#ggplot(df_360_meta %>% filter(id %in% c('PBS')), aes(x = dna, y = intensity, color = id)) + geom_jitter(shape = 21)


```

## 485nm ex

```{r}
ggplot(df_485_meta %>% filter(id %in% c('control','empty','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity)) + geom_point() + facet_wrap(~id, scales = 'free')

ggplot(df_485_meta %>% filter(id %in% c('control','empty','PBS','PCA','PCN','PYO')), aes(x = dna, y = intensity, color = id)) + geom_point() + geom_smooth(method = 'lm')
```

## Figure 

```{r}

df_fig <- bind_rows(df_360_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PYO','PBS','control')) %>% mutate(ex = '360'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCA','PBS','control')) %>% mutate(ex = '485')
          )

df_fig <- df_fig %>% group_by(id,ex) %>% mutate(max = max(mean)) %>% mutate(norm_int = intensity / max, norm_mean = mean / max)

#ggplot(df_fig, aes(x = dna, y = norm_int, color = id, shape = ex)) + geom_point()

#ggplot(df_fig, aes(x = dna, y = norm_mean, color = id, shape = ex)) + geom_point()

#ggplot(df_fig, aes(x = dna, y = norm_mean, color = id, shape = ex, linetype = ex)) + geom_point(aes(y = norm_int)) + geom_line() + ylim(0,NA)

ggplot(df_fig, aes(x = dna, y = norm_mean, color = id)) + geom_point(aes(y = norm_int)) + geom_line() + ylim(0,NA) + facet_wrap(~ex)

#ggplot(df_fig %>% filter(id != 'PBS'), aes(x = dna, y = norm_mean, color = id)) + geom_jitter(aes(y = norm_int)) + geom_line() + ylim(0,NA)

facet_names <- list(
  '360'="PYO",
  '485'="PCA"
)

facet_labeller <- function(variable,value){
  return(facet_names[value])
}

ggplot(df_fig %>% filter(id != 'PBS'), aes(x = dna, y = norm_mean, color = id)) + geom_jitter(aes(y = norm_int), shape = 21) + geom_line() + 
  facet_wrap(~ex, labeller = facet_labeller, scales = 'free') + 
  scale_color_manual(values = c('light gray','black','black')) + scale_y_continuous(labels = scales::percent, limits = c(0,NA)) + 
  labs(x = 'bp DNA (uM)', y = 'Normalized fluorescence')
```

Trying to make figure to match etbr and mst data:

```{r}
df_fig_2 <- bind_rows(df_360_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PYO','control')) %>% mutate(ex = '360_PYO'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCA','control')) %>% mutate(ex = '485_PCA'),
          df_485_meta %>% group_by(id, dna) %>% mutate(mean = mean(intensity)) %>% filter(id %in% c('PCN','PBS')) %>% mutate(ex = '485_PCN')
          )

df_fig_2 <- df_fig_2 %>% group_by(id,ex) %>% mutate(max = max(mean)) %>% mutate(norm_int = intensity / max, norm_mean = mean / max) %>% 
  ungroup() %>% 
  mutate(ex = fct_relevel(ex,'485_PCA','485_PCN','360_PYO' ))

facet_names <- list(
    '485_PCA'="PCA",'485_PCN'="PCN",'360_PYO'="PYO"
)

facet_labeller <- function(variable,value){
  return(facet_names[value])
}

#df_fig_2$ex = factor(df_fig_2$ex, levels=c('485_PCA','485_PCN','360_PYO'))

#scale_color_manual(values = c('light gray','black','black')) + 

ggplot(df_fig_2, aes(x = dna, y = norm_mean, color = id)) + geom_jitter(aes(y = norm_int), shape = 21) + geom_line() + 
  facet_wrap(~ex, scales = 'free', ncol = 1) + 
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = 'bp DNA (uM)', y = 'Normalized fluorescence') + scale_color_manual(values = c('light gray','light gray','black','black','black')) + guides(color = F) + xlim(0,300)

```
Conclusions: does not seem to be signficant fluorescence change compared to empty wells. Note that there is nonzero variation between wells in plate reader. Reduced phenazines do not bind strongly to DNA.

Caveats: 90uM phenazine may be well below the Kd, so this only provides some evidence that there's no binding at this concentration of phenazine.

```{r}
sessionInfo()
```