---
title: "PYO FRAP"
subtitle: "Analysis"
fontsize: 12pt
date: "05_30_19 "
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")


source("../../IDA/tools/plotting_tools.R")

theme_set(theme_1())
```

# Intro

Some motivation about why we're doing FRAP etc. 


# Methods

How were the data acquired etc. 


# Analysis

First, let's read in the csv of the frap data. All of these scans were acquired in the same way, so they should be pretty comparable. In excel I already normalized the intensity values to the average baseline intensity (before bleach).

```{r}

df <- read_csv("05_30_19_PYO_FRAP_data.csv")

df
```

Now let's look at some overview plots.

```{r}
ggplot(df %>% filter(scaled > 0.1), aes(x = time, y = scaled, color = condition)) + geom_point(shape = 21) 

ggplot(df %>% filter(scaled > 0.1), aes(x = time, y = scaled, color = condition, group = id)) + geom_line(alpha = 0.5) + facet_wrap(~condition)
```

```{r}
  
avg_max <- df %>% 
  filter(img_num >= 240) %>% 
  group_by(condition,id) %>% 
  mutate(avg_max = mean(intensity)) %>% 
  select(condition, id, avg_max) %>% 
  distinct()

df_norm <- left_join(df, avg_max, by = c('condition', 'id')) %>% 
  filter(intensity > 1) %>% 
  group_by(condition, id) %>% 
  mutate(min = min(intensity)) %>% 
  mutate(bg_intensity = intensity - min) %>% 
  mutate(bg_avg_max = avg_max - min) %>% 
  mutate(bg_scaled = bg_intensity / bg_avg_max)

ggplot(df_norm, aes(x = time, y = bg_scaled, color = condition, group = id)) + geom_line(alpha = 0.5) + facet_wrap(~condition)

ggplot(df_norm %>% filter(condition=='1pctAgaroseDNA'), aes(x = time, y = bg_scaled, color = factor(id))) + geom_line()

ggplot(df_norm, aes(x = time, y = bg_scaled, color = condition)) + geom_point(shape = 21) + geom_hline(yintercept = 0.5)+scale_color_viridis_d()
```

```{r}
df_thresh <- df_norm %>% 
  group_by(condition,id) %>% 
  filter(bg_scaled > 0.5 & time > 5) %>% 
  filter(time == min(time)) %>% 
  mutate(half_time = time) %>% 
  select(condition, id, half_time)

df_bleach <- df_norm %>% 
  group_by(condition,id) %>% 
  filter(time > 5) %>% 
  filter(time == min(time)) %>% 
  mutate(bleach_time = time) %>% 
  select(condition, id, bleach_time)

df_tau <- left_join(df_thresh, df_bleach, by = c('condition', 'id')) %>% 
  mutate(tau = half_time - bleach_time) %>% 
  group_by(condition) %>% 
  mutate(mean = mean(tau))

ggplot(df_thresh, aes(x = condition, y = half_time)) + geom_point() + geom_point(data = df_bleach, aes(x = condition, y = bleach_time), color = 'red')

ggplot(df_tau, aes(x = condition, y = tau)) + geom_point() + geom_point(aes(y = mean), color = 'red')
```

```{r}
soumpasis_fit <- function(t_d, t){
  exp(1)^((-2*t_d) / t) * (besselI(t,0,expon.scaled = T) + besselI(t,1,expon.scaled = T))
}

df_frap = tibble(frap = soumpasis_fit(t_d = 5, t = seq(1,25, 0.1)), t = seq(1,25, 0.1))

ggplot(df_frap, aes(x = t, y = frap )) + geom_point()
```

```{r}
df_4pct_1 <- df_norm %>% filter(condition == '4pctAgarose' & id ==1) %>% filter(time>5) %>% mutate(time = time - 6.2)

ggplot(df_4pct_1, aes(x = time, y = bg_scaled)) + geom_point() + 
    geom_smooth(method='nls',formula=y~(I0 - a*exp(-B * x)), method.args=list(start=c(I0 = 0.85, a = 0.5, B = 0.563)),se=F)
                

nls(formula = bg_scaled~I0 - a*exp(-B * time), data = df_4pct_1, start = list(I0 = 0.85, a = 0.5, B = 0.563) )
                
log(2) / 0.14 
```