---
title: "Propidium and Ethidium with Colony Biofilms"
subtitle: "LC-MS Quantifications from exp 07-23-19"
author: "Scott Saunders"
fontsize: 12pt
date: '09-23-2019'
output:
  html_document:
    theme: cosmo
    code_folding: hide
    self_contained: no
    toc: yes
  pdf_document:
    toc: yes
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(viridis)
library(knitr)
library(kableExtra)



knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../../../IDA/tools/plotting_tools.R")

theme_set(theme_1())
```

# Methods

## Samples

### WT

WT colonies for TOTO measurement

* WT Toto - 6 biofilm, 6 agar

WT controls for EtBr

* WT H2O - 3 cells, 3 agar

WT controls for PI

* WT dmso - 3 cells, 3 agar

WT exp w/ EtBr

* WT EtBr 10uM  - 3 cells, 3 agar

* WT EtBr 100uM  - 3 cells, 3 agar

* WT EtBr 500uM  - 3 cells, 3 agar

WT exp w/ PI

* WT PI 10uM  - 3 cells, 3 agar

* WT PI 100uM  - 3 cells, 3 agar

* WT PI 200uM  - 3 cells, 3 agar

 = 60 samples
 
------

### ∆phz*

∆phz* + PYO + EtBr 

* ∆phz* + 50uM PYO + 10uM EtBr - 2 cells, 2 agar

* ∆phz* + 50uM PYO + 100uM EtBr - 2 cells, 2 agar

* ∆phz* + 50uM PYO + 500uM EtBr - 2 cells, 2 agar 

∆phz* + PCN + EtBr 

* ∆phz* + 50uM PCN + 10uM EtBr - 2 cells, 2 agar

* ∆phz* + 50uM PCN + 100uM EtBr - 2 cells, 2 agar

* ∆phz* + 50uM PCN + 500uM EtBr - 2 cells, 2 agar 

∆phz* + PYO + PI 

* ∆phz* + 50uM PYO + 10uM PI - 2 cells, 2 agar

* ∆phz* + 50uM PYO + 100uM PI - 2 cells, 2 agar

* ∆phz* + 50uM PYO + 200uM PI - 2 cells, 2 agar 

∆phz* + PCN + PI 

* ∆phz* + 50uM PCN + 10uM PI - 2 cells, 2 agar

* ∆phz* + 50uM PCN + 100uM PI - 2 cells, 2 agar

* ∆phz* + 50uM PCN + 200uM PI - 2 cells, 2 agar 

 = 48 samples
 
 --------
 
  = 108 samples total
  
  First run is WT H2O and WT WT EtBr Samples
  
  
  
  
  
  
  ------
  
```{r}
sessionInfo()  
```
