---
title: "Figure 4 - PHZ eDNA Paper"
fontsize: 12pt
date: "11_05_18"
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---

**Note** the YAML contains specifications for a github document and html. The best way to deal with this is to knit them separately from the knit menu. Otherwise the html has blurry plots and tends to delete the md cached plots unless you tell it to cache everything!

```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)

knitr::opts_knit$set(root.dir = '/Users/scottsaunders/git/labwork/Figures/draft1/')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

theme_1 <- function () { 
    theme_classic() %+replace% 
        theme(
          axis.text = element_text( size=12),
          axis.title=element_text(size=14),
          strip.text = element_text(size = 14),
          strip.background = element_rect(color='white'),
          legend.title=element_text(size=14),
          legend.text=element_text(size=12),
          panel.grid.major = element_line(color='grey',size=0.1)
        )
}

theme_set(theme_1())
```


The SWV decay and Dap estimation figure is stored in DapEstimation notebook and the dPHZ vs. blank IDA notebook. Need to work on adapt everything for reproducing here. 

```{r estimated Dap}

df_dap <- tibble(material=c('blank','biofilm','blank','biofilm'),
       Dap=c(1e-5,1e-6,2.8e-6,1.8e-8),
       Dcat=c('electron','electron','PYO','PYO'))
ggplot(df_dap,aes(x=Dcat,y=Dap,color=Dcat))+
  geom_point()+facet_wrap(~material,scales='free')+
  scale_y_log10(limits=c(1e-9,1e-5))+
  guides(color=F)
```


```{r DNACT no O2}
df_noO2 <- read_csv('../data/DNA_modified_electrode_noO2.csv',comment = "#") %>% 
  mutate(Construct=fct_relevel(Construct,'wm','mm'))

dnaCT_noO2 <- ggplot(df_noO2,aes(x=E,y=Current,color=Construct))+
  geom_path(size=1.5,lineend = 'round')+
  scale_x_reverse()+
  scale_color_viridis_d(breaks = c('wm','mm'), labels=c("Well matched DNA","Mismatched DNA")) +
  labs(x='E (V vs. Ag/AgCl)',y='Current (Amps)')+
  theme(legend.position = c(0.3, 0.8),
        legend.title = element_blank(),
        legend.background = element_rect(fill=NA))

dnaCT_noO2
```

```{r DNACT w O2}
df_wO2 <- read_csv('../data/DNA_modified_electrode_wO2.csv',comment = "#") %>% 
  mutate(Condition=paste(Construct,Aerobic,PCN,sep='_')) %>% 
  mutate(Condition=fct_relevel(Condition,"wm_TRUE_TRUE","mm_TRUE_TRUE","wm_TRUE_FALSE"))


dnaCT_wO2 <- ggplot(df_wO2 %>% filter(Aerobic==T),aes(x=E,y=Current,color=Condition))+
  geom_path(size=1.5,lineend = 'round')+
  scale_x_reverse()+
  scale_color_viridis_d(
    breaks = c('wm_TRUE_TRUE','mm_TRUE_TRUE','wm_TRUE_FALSE'), 
    labels=c(expression("wmDNA"+"PCN"+O[2]),expression("mmDNA"+"PCN"+O[2]),expression("wmDNA"-"PCN"+O[2]))
    ) +
  labs(x='E (V vs. Ag/AgCl)',y='Current (Amps)')+
  theme(
    legend.position = c(0.3, 0.8),
    legend.title = element_blank(),
    legend.background = element_rect(fill=NA))

dnaCT_wO2
```

```{r}
dnaCT <- plot_grid(dnaCT_noO2,dnaCT_wO2,ncol=2,labels=c("E","F"),align='hv',axis='tblr',scale=0.9)

dnaCT

ggsave('fig4ef_dnaCT.pdf',dnaCT,height = 4 , width = 10)

```