---
title: "SWV and GC calibration - Blank"
fontsize: 12pt
date: "02_07_19 "
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)
library(knitr)
library(kableExtra)
library(patchwork)


knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../../IDA/tools/echem_processing_tools.R")
source("../../../IDA/tools/plotting_tools.R")

theme_set(theme_1())
```

# Intro

Let's try to settle on an estimate for $n$ and $\psi$ and by looking for consistent $D_{ap}$ estimates from the SWV and GC blank datasets with known concentrations. We can extract a $D_{ap}$ by comparing $I_{GC}$ to $I_{SWV}$, but because we know the bulk concentration for this in vitro system we can also estimate $D_{ap}$ by comparing $I_{GC}$ to $C$, or $I_{SWV}$ to $C$. 


# SWV Fast and GC Dataset

Let's first read in the processed fast dataset.
```{r}
df_blank <- read_csv("../data/11_28_18_swvFast_gc_soak_processed.csv") 
```

We'll select just the subset from SWV electrode 1 and GC electrode 2 (collector). Let's also transform the added PYO from $\mu M$ to $\frac{mol}{cm^3}$. 

```{r}
df_cal <- df_blank %>% 
  filter(electrode_from_swv=='i1' & electrode_from_gc=='i2') %>% 
  mutate(C_mol_cm3 = PHZaddedInt*(10^-9))

```

## SWV vs. GC

The data look like this:
```{r}
plot_swvGC <- ggplot(df_cal, aes(x = signal_from_swv, y = signal_from_gc)) + 
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_swvGC_styled <- plot_swvGC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(I[swv]~(nA)), y = expression(I[gc]~(nA)))
  
plot_swvGC_styled
```

Recall that for this type of data:

$$D_{ap} = \pi t_p \left( \frac{m A \psi}{S} \right) ^2$$    
So, here's a function that will calculate $D_{ap}$ from the slope of this data:
```{r}
dap_from_swvGC <- function(m, t_p=1/(2*300)){
  
  psi <-  0.7
  #psi <-  0.75
  #A <-  0.013 #cm^2
  A <-  0.025 #cm^2
  S <-  18.4 #cm
  d_ap <- (m*A*psi)^2 / (S^2 * pi * t_p)
  
  d_ap
}
```

And now, we'll fit a linear model and calculate a $D_{ap}$.
```{r}
lm_swvGC <- tidy(lm(signal_from_gc~signal_from_swv, data = df_cal), conf.int = T) %>% 
  filter(term=='signal_from_swv') %>% 
  mutate(dap=dap_from_swvGC(m = estimate)) %>% 
  mutate(dap_high = dap_from_swvGC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_swvGC(m = conf.low)) %>% 
  mutate(dataset = 'SWVvsGC')

glance(lm(signal_from_gc~signal_from_swv, data = df_cal))


#write_csv(lm_swvGC,"../data/02_07_19_blank_dap_estimates.csv")

lm_swvGC %>% 
  kable() %>% 
  kable_styling()
```
Ok, seems reasonable, let's move forward and we'll compare at the end. 

## SWV

Here's the data:
```{r}
plot_SWV <- ggplot(df_cal, aes(x = C_mol_cm3, y = signal_from_swv)) +
    geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_SWV_styled <- plot_SWV+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[swv]~(nA)))
  
plot_SWV_styled
```

For a plot of $I_{SWV}$ vs. $C$, the slope can be definited in terms of the parameters for $I_{SWV}$ other than concentration. This yields an expression for $D_{ap}$:

$$D_{ap} = \pi t_p \left( \frac{m}{\psi n F A} \right) ^2 $$
So we'll again define the function to calculate it:
```{r}
dap_from_SWV <- function(m, t_p=1/(2*300)){
  
  n=1.8
  #n=2
  psi <-  0.7
  #psi <-  0.75
  
  #A <-  0.013 #cm^2
  A <-  0.025 #cm^2
  
  F_C = 96485.332
  
  d_ap <- pi * t_p * (m / (psi * n * F_C * A))^2
  
  d_ap
}
```

The linear model of this data gives the following:

```{r}
lm_SWV <- tidy(lm(signal_from_swv~C_mol_cm3, data = df_cal), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_SWV(m = estimate)) %>% 
  mutate(dap_high = dap_from_SWV(m = conf.high)) %>% 
  mutate(dap_low = dap_from_SWV(m = conf.low)) %>% 
  mutate(dataset = 'SWV')

glance(lm(signal_from_swv~C_mol_cm3, data = df_cal))

lm_SWV %>% 
  kable() %>% 
  kable_styling()
```
Also seems reasonable. 

## GC

Here's the GC vs. Concentration data:
```{r}
plot_GC <- ggplot(df_cal, aes(x = C_mol_cm3, y = signal_from_gc)) +    
  geom_smooth(method='lm', linetype = 2, color = 'black') + geom_point(shape = 21)

plot_GC_styled <- plot_GC+  
  scale_x_continuous(labels = nA_label) + 
  scale_y_continuous(labels = nA_label) + 
  labs(x = expression(PYO~(mu*M)), y = expression(I[gc]~(nA)))
  
plot_GC_styled
```
Again, very linear.

$I_{GC}$ is simply $I = nFSDC$, so $D_{ap}$ is:

$$D_{ap} = \frac{m}{n F S}$$

And here's our function.

```{r}
dap_from_GC <- function(m){
  
  n=1.8
  #n=2
  
  S <-  18.4 #cm
  
  F_C = 96485.332
  
  d_ap <- m / (n * F_C * S)
  
  d_ap
}
```

The linear model:
```{r}
lm_GC <- tidy(lm(signal_from_gc~C_mol_cm3, data = df_cal), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_GC(m = estimate)) %>% 
  mutate(dap_high = dap_from_GC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_GC(m = conf.low)) %>% 
  mutate(dataset = 'GC')

glance(lm(signal_from_gc~C_mol_cm3, data = df_cal))

lm_GC %>% 
  kable() %>% 
  kable_styling()
```
## Comparison

Now let's look at the estimates from each subset of the data:
```{r}
lms <- bind_rows(lm_swvGC,lm_SWV,lm_GC) %>% 
  select(dataset, term, dap, dap_low, dap_high)

lms %>% 
  kable() %>% 
  kable_styling()
```
So, you can see that the estimates actually match very closely!!

Here it is plotted:
```{r}
plot_cal_dap <- ggplot(lms, aes(x = dataset, y = dap)) + 
  geom_pointrange(aes(ymin = dap_low, ymax = dap_high)) + 
  ylim(0, 1e-5)

plot_cal_dap_styled <- plot_cal_dap+
  labs(x = 'Dataset', y = expression(D[ap]~(cm^2 / sec)))+
  scale_x_discrete(breaks = c('GC','SWV', 'SWVvsGC'), labels=c('GC vs.\n[PYO]', 'SWV vs.\n[PYO]', 'GC vs.\nSWV'))

plot_cal_dap_styled
```


```{r}
cal_dap_grid <- plot_grid(plot_GC_styled, plot_SWV_styled, plot_swvGC_styled, plot_cal_dap_styled, ncol = 2, align = 'hv', axis = 'tblr', scale = 0.9, labels = c("A","B","C","D"))

cal_dap_grid

save_plot("../../supplement/figures/03_22_19_fig_supp_dapCal.pdf",cal_dap_grid, base_width = 6.5, base_height = 5)
```
So, you can see that the estimates correspond very nicely when we use these parameters. Realizing that these estimates should match led me to look for a better estimate of $\psi$ based on our scan parameters and I found this [reference](https://link.springer.com/book/10.1007%2F978-3-642-02915-8), which suggested that for $n = 2$ and $E_{swv} = 25mV, \psi = 0.75$ and for $n = 1.8, \psi = 0.7$. I tested both and the second parameter set yielded good results, and corresponds to the peak splitting we see in CVs (n is not quite 2). Therefore the parameters I would like to proceed with are:

* $A = 0.025 \text{cm}^2$ as measured by Fernanda  
* $S = 18.4 \text{cm}$  
* $n = 1.8$ 
* $\psi = 0.7$ 
* $t_p = \frac{1}{600}$ for SWV fast settings (one half the SWV period, which is 1/freq)

Of course the first thing I did was go back to the aggregate biofilm dataset and use these parameters to recalculate the biofilm values. Interestingly, many of the values lie higher than the blank IDA estimate. I also saw that the slow SWV blank dataset gave a different value, and that confused me. So I wanted to evaluate it here. 


# SWV Slow and GC Dataset

Here I import the dataset, and calculate $D_{ap}$ values from each subset of the data (using $t_p = \frac{1}{30}$). I also defined some alternative functions to play around with the parameters, you can ignore those for now. Here I am using the exact same functions from the SWV fast dataset above. 

```{r}

df_slow <- read_csv("../data/11_28_18_swvSlow_gc_soak_processed.csv") 

df_slowCal <- df_slow %>% 
  filter(electrode_from_swv=='i1' & electrode_from_gc=='i2') %>% 
  mutate(C_mol_cm3 = PHZaddedInt*(10^-9))

dap_from_SWV2 <- function(m, t_p=1/(2*300)){
  
  #n=1.8
  n=2
  #psi <-  0.7
  psi <-  0.75
  
  #A <-  0.013 #cm^2
  A <-  0.025 #cm^2
  
  F_C = 96485.332
  
  d_ap <- pi * t_p * (m / (psi * n * F_C * A))^2
  
  d_ap
}

dap_from_swvGC2 <- function(m, t_p=1/(2*300)){
  
  #psi <-  0.7
  psi <-  0.75
  #A <-  0.013 #cm^2
  A <-  0.025 #cm^2
  S <-  18.4 #cm
  d_ap <- (m*A*psi)^2 / (S^2 * pi * t_p)
  
  d_ap
}

lm_swvGC_slow <- tidy(lm(signal_from_gc~signal_from_swv, data = df_slowCal), conf.int = T) %>% 
  filter(term=='signal_from_swv') %>% 
  mutate(dap=dap_from_swvGC(m = estimate, t_p = (1/30))) %>% 
  mutate(dap_high = dap_from_swvGC(m = conf.high, t_p = (1/30))) %>% 
  mutate(dap_low = dap_from_swvGC(m = conf.low, t_p = (1/30))) %>% 
  mutate(dataset = 'SWVvsGC')

lm_SWV_slow <- tidy(lm(signal_from_swv~C_mol_cm3, data = df_slowCal), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_SWV(m = estimate, t_p = (1/30))) %>% 
  mutate(dap_high = dap_from_SWV(m = conf.high, t_p = (1/30))) %>% 
  mutate(dap_low = dap_from_SWV(m = conf.low, t_p = (1/30))) %>% 
  mutate(dataset = 'SWV')


lm_GC_slow <- tidy(lm(signal_from_gc~C_mol_cm3, data = df_slowCal), conf.int = T) %>% 
  filter(term=='C_mol_cm3') %>% 
  mutate(dap=dap_from_GC(m = estimate)) %>% 
  mutate(dap_high = dap_from_GC(m = conf.high)) %>% 
  mutate(dap_low = dap_from_GC(m = conf.low)) %>% 
  mutate(dataset = 'GC')

lms_slow <- bind_rows(lm_swvGC_slow,lm_SWV_slow,lm_GC_slow) %>% 
  select(dataset, term, dap, dap_low, dap_high)

lms_slow %>% 
  kable() %>% 
  kable_styling()
```
Ok, so obviously the GC is still the same, because we didn't do anything different. However, The SWV and SWVvsGC estimates of $D_{ap}$ are now wildly different!? The SWV estimate is too high and the SWV vs. GC estimate is too low. This suggests that $n$ and $\psi$ could be too low. I tried changing the values to 2 and 0.75, but the change was minimal. Therefore, I do not know why these estimates are so off, but I would definitely trust the SWV fast datasets more until we understand why. 


With psi = 0.75 and n = 2:

* SWV = 1.33e-5 
* SWV v GC = 2.93e-6 

With psi = 0.7 and n = 1.8:

* SWV = 1.897 e-5
* SWV v GC = 2.55e-6

And just for clarity, here you can see that the plots are still extremely linear.
```{r}
p1 <- ggplot(df_slowCal, aes(x = signal_from_swv, y = signal_from_gc)) + geom_point() + geom_smooth(method='lm')

p2 <- ggplot(df_slowCal, aes(x = C_mol_cm3, y = signal_from_swv)) + geom_point() + geom_smooth(method='lm')

p3 <- ggplot(df_slowCal, aes(x = C_mol_cm3, y = signal_from_gc)) + geom_point() + geom_smooth(method='lm')

p1 | p2 | p3
```

# Conclusions

I am glad that I did this, but it is also unsettling that changing the scan rate throws off the estimate drastically even with a blank electrode. For now, let's move forward with the parameters I used in the SWV fast section above. 


*****
