---
title: "Figure 1 - Draft 2"
fontsize: 12pt
date: "02_05_19"
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---

**Note** the YAML contains specifications for a github document and html. The best way to deal with this is to knit them separately from the knit menu. Otherwise the html has blurry plots and tends to delete the md cached plots unless you tell it to cache everything!

```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(kableExtra)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

theme_1 <- function () { 
    theme_classic() %+replace% 
        theme(
          axis.text = element_text( size=12),
          axis.title=element_text(size=14),
          strip.text = element_text(size = 14),
          strip.background = element_rect(color='white'),
          legend.title=element_text(size=14),
          legend.text=element_text(size=12),
          panel.grid.major = element_line(color='grey',size=0.1)
        )
}
theme_set(theme_1())
```

# Panel D

Read in the data - concentrations quantified from LC (Absorbance) peaks with standard curve.
```{r}
wt_extracts <- read_csv('../data/HPLC_data_colonies_02_16_18.csv')

wt_extracts %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```


Now we need to convert the concentration (Amount column) into the concentration that existed in the biofilm or the agar, not the sample tube. For the agar, we know that there is 5mL and in this case 2mL of solution was added for extraction. For the biofilm, we estimate that the volume of the colony is about 60uL, and we resuspended it in 800uL. Here, I'll also calculate the means, so we can look at the mean and individual measurements.
```{r}
# Calculate concentration from Amount. For biofilm: Amount * 800 / 62, for agar: Amount * 7/5
wt_conc <- wt_extracts %>% 
  mutate(phz_conc = ifelse(material=='biofilm', Amount * (800 / 62), Amount * (7 / 5)))

# Calculate the mean concentration for each group - 'Name' is the measured phenazine
# Only assign mean to first observation, because we only need one bar
wt_conc_means <- wt_conc %>% 
  group_by(material, Name) %>% 
  mutate(mean = ifelse(replicate==1, mean(phz_conc), NA)) 
```

Let's plot the data with a fixed scale, so we can see the pattern of phenazine production by the WT cells:
```{r}
#Plot layout
plot_wt_fixed <- ggplot(wt_conc_means, aes(x = material, fill = material)) + 
  geom_col(aes(y = mean)) + 
  geom_jitter(aes(y = phz_conc), height = 0, width = 0.1, shape = 21) +
  facet_wrap(~Name)

#Plot styling
plot_wt_fixed +
  labs(x = NULL, y = expression("Phenazine concentration" ~ ( mu*M ))) + 
  theme(axis.title.x = element_text(size = 14)) + 
  scale_fill_manual(guide = F, values = c("#F0E442","light green"))
         
```

Here you can see that the abundance of phena

```{r}
#Plot layout
plot_wt_free <- ggplot(wt_conc_means, aes(x = material, fill = material)) + 
  geom_col(aes(y = mean)) + 
  geom_jitter(aes(y = phz_conc), height = 0, width = 0.1, shape = 21) +
  facet_wrap(~Name, scales = 'free')

#Plot styling
plot_wt_free_styled <- plot_wt_free +
  labs(x = NULL, y = expression("Phenazine concentration" ~ ( mu*M ))) + 
  theme(axis.title.x = element_text(size = 14)) + 
  scale_fill_manual(guide = F, values = c("#F0E442","light green"))
    
plot_wt_free_styled     
```
ok and now this makes our point.

# Panel E

Let's first read in the data. This csv contains the data for multiple experiments that were run on the LC-MS on the same day, so let's just look at the dPHZ* colonies.

```{r}
# Read in only the dPHZstar data, and convert the amount of phenazine added to a numeric
dphz_extracts <- read_csv('../data/2018_10_30_HPLC_concentrations_df.csv',comment = "#") %>% 
  filter(strain == 'dPHZstar') %>% 
  mutate(added_phz_num=as.double(str_extract(amount_added,"^[0-9.]+")))

dphz_extracts %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>% 
  scroll_box(height = '250px')
```

You can see that I've already calculated the biofilm concentrations for this dataset. You can also see that each colony was recorded with the amount of synthetic phenazine added to the agar. Let's calculate means for each group and look at an overview of the measurements.

```{r}
# Names for facet panels
facet_labels = c(PCA = "PCA added", PCN = "PCN added", PYO = "PYO added")

# Plot layout
plot_dphz_overview <- ggplot(dphz_extracts, aes(x = measured_phenazine, y = calcConc, fill = amount_added)) + 
  geom_jitter(shape = 21, height = 0, width = 0.1, size = 2)+
  facet_wrap(~added_phenazine, scale = 'free', labeller = labeller(added_phenazine = facet_labels))

# Plot styling
plot_dphz_overview +
  scale_fill_viridis(discrete = T, breaks = c('0.1uM','1uM','10uM','50uM','100uM','200uM')) +
  labs(fill = "Amount of\nphenazine added" , 
       x = "Phenazine measured", 
       y = expression("Biofilm concentration" ~ ( mu*M )) )
```

Ok, so we can see that when we add a phenazine to this strain, we only measure significant amounts of that phenazine. Therefore, this mutant is "phenazine inert." There is a little bit of PCN detected, but upon inspection of the chromatograms and MS, it is a persistent background peak that is not actually PCN.

So, let's proceed and just look at the measured phenazines that match the added phenazine, and we'll calculate some means.

```{r}
# Calculate Means
dphz_extracts_means <- dphz_extracts %>% 
  filter(measured_phenazine==added_phenazine) %>% 
  group_by(measured_phenazine,amount_added) %>% 
  mutate(mean = mean(calcConc))


# Plot layout
plot_dphz_binding <- ggplot(dphz_extracts_means, aes(x=added_phz_num,y=calcConc))+
  geom_line(size=1, aes(y=mean,color=measured_phenazine))+
  geom_point(size=2, shape = 21, aes(fill=measured_phenazine))

# Plot styling
plot_dphz_binding_styled <- plot_dphz_binding + 
  labs(x=expression('Phenazine in agar ('~mu*"M)"), 
       y = expression("Biofilm concentration" ~ ( mu*M )) ) +
  theme(legend.position = c(0.5, 0.9),
        legend.title = element_blank(),
        legend.background = element_rect(fill=NA))


#coord_fixed(ratio = 0.75)

plot_dphz_binding_styled
```

# Combine Panels 
```{r}
fig_1 <- plot_grid(plot_wt_free_styled, plot_dphz_binding_styled, align = 'hv', axis = 'tblr', rel_widths = c(1.5,1), aspec)

fig_1
```

```{r eval = F}

save_plot("../figures/02_05_19_fig_1_draft_2.pdf", fig_1, base_aspect_ratio = 2)

```

