---
title: "Inter phenazine electron transfer +/- ctDNA"
author: "Scott Saunders"
fontsize: 12pt
date: '10_18_19'
output:
  html_document:
    theme: cosmo
    code_folding: hide
    self_contained: no
    toc: yes
  github_document:
    pandoc_args: --webtex
---

```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(viridis)
library(knitr)
library(kableExtra)



knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../IDA/tools/plotting_tools.R")

theme_set(theme_1())
```

# Intro

This experiment was an extension of the [09_17_2019 experiment](https://scott-saunders.github.io/labwork/Spectroscopy/2019_09_17_solution_ET/2019_09_17_solution_ET_analysis_2.html) in which I clearly established inter phenazine ET in solution. Upon starting to write that section of the paper (in vitro ET), I thought it may be useful to test whether or not DNA inhibited the reaction, since PYOox binds strongly to DNA. I assumed DNA would not inhibit, since it's an equilibrium and there's always some unbound...but I thought it was worth testing and pretty straightforward. We also knew that PodA could function on PYO in the presence of DNA.


# Methods

Experiments were performed in a similar manner to the previous inter-phenazine solution ET notebook (09_17_2019). Mixtures were prepared in the anaerobic chamber with electrochemically reduced phenazines in serum bottles and oxidized phenazine solutions that had been in the chamber for at least 24 hrs. Measurements were done in black walled clear bottom plates that had been in the chamber > 1 week. 

Each well contained the following.

* 67.5uL PYOox (100uM) or PBS
* 67.5uL PHZred (100uM) or PBS
* 15uL ctDNA (2mg/mL) or PBS

150uL per well.

The plate reader program was - abs spectrum, kinetic run (~2min) with abs690, em460, em528, then another abs spectrum. Rows were setup one at a time with 6 wells each. Columns 1-3 did not contain DNA (PBS instead) and columns 4-6 did contain ctDNA.

* Row A: PYOox baseline
* Row B: PYOred baseline
* Row C: PCAred and PYOox
* Row D: PCNred and PYOox

# Results

Here's the info for each well:
```{r}
metadata <- read_csv("data/2019_10_18_ET_ctDNA_metadata.csv")

metadata %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

## Abs Spectrum 1

Now let's read in the first absorbance spectra taken for each condition and we'll add the appropriate metadata:
```{r}

df <- read_csv("data/2019_10_18_ET_ctDNA_abs_spectra_1.csv") %>%
    gather(key = 'well', value = 'intensity', -Wavelength)

df_meta <- left_join(df, metadata, by = 'well')

df_meta %>% kable %>% kable_styling() %>% scroll_box(height = '400px')

```

Let's look at all the data:

```{r}
#ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_smooth(se = T)

ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + geom_point()

```

This looks pretty much like the past notebook with interphenazine ET. If we look out at 690 for the PYO abs peak it looks like the orange PYOox is higher than all the other conditions. Let's zoom into that region for a closer look:

```{r}
ggplot(df_meta, aes(x = Wavelength, y = intensity, color = red_ox)) + 
  geom_jitter(shape=21)+coord_cartesian(xlim = c(600,700), ylim = c(0,0.2))
```

Ok, this looks pretty much as we expected. All of the conditions show 690 abs closer to PYOred than PYOox.

Let's try and condense this info into some pointranges:

```{r}
df_690 <- df_meta %>% group_by(red_ox, Wavelength, DNA) %>% mutate(mean = mean(intensity), sd = sd(intensity))  %>% filter(Wavelength == 690)

ggplot(df_690, aes(x = red_ox, y = intensity, color = DNA )) + 
  geom_jitter(shape = 21, width = 0.1) + 
  geom_pointrange(data = . %>% filter(rep ==1), 
                  aes(y = mean,ymin = mean-2*sd,ymax = mean+2*sd), 
                  position = position_jitter(width = 0.1))
```

Ok these pointranges show mean +/- 2 standard deviations. So it looks like for all of the conditions there's no significant difference +/- ctDNA.

## Abs690 Kinetic

The Spectrum data looks good, let's also take a look at the kinetic data at 690 to see if there's a difference in the reaction rates. First we'll read it in:

```{r}

df_kin <- read_csv("data/2019_10_18_ET_ctDNA_abs_690kinetic_1.csv") %>%
    gather(key = 'well', value = 'intensity', -min)

df_kin_meta <- left_join(df_kin, metadata, by = 'well')

df_kin_meta %>% kable() %>% kable_styling() %>% scroll_box(height = '400px')

```

Now let's look at the values over time (~2min)

```{r}

ggplot(df_kin_meta, aes(x = min, y = intensity, color = DNA)) + geom_point() + facet_wrap(~red_ox)

```

Ok, the intensity trends we observed with the spectra looks the same here. There are no obvious differences in slope, so I still don't think we can say anything about reaction rates...they all proceed too fast.

Let's collapse this data to pointranges too and see how it looks:

```{r}
df_kin_well <- df_kin_meta %>% 
  group_by(red_ox, DNA, rep) %>% 
  summarise(well_mean = mean(intensity), well_sd = sd(intensity))

ggplot(df_kin_well, aes(x = red_ox, y = well_mean, color = DNA)) + 
  geom_pointrange(aes(ymin = well_mean - 2*well_sd, ymax = well_mean + 2*well_sd), 
                  position = position_jitter(width = 0.1))
```

As expected the variation across the kinetic timepoints is very small. Let's aggregate further into single point ranges percondition:

```{r}
df_kin_exp <- df_kin_well %>% 
  group_by(red_ox, DNA) %>% 
  summarise(exp_mean = mean(well_mean), exp_sd = sd(well_mean))

ggplot(df_kin_exp, aes(x = red_ox, y = exp_mean, color = DNA)) + 
  geom_pointrange(aes(ymin = exp_mean - 2*exp_sd, ymax = exp_mean + 2*exp_sd), 
                  position = position_dodge(width = 0.2))
```

### Figure

Looks good, now let's make a pretty figure:

```{r}
labels = labels = c('PBS + PYO[ox]','PYO[red] + PBS','PCA[red] + PYO[ox]','PCN[red] + PYO[ox]')


plot_ctDNA_690 <- ggplot(df_kin_exp, aes(x = red_ox, y = exp_mean, color = DNA)) +  
  geom_pointrange(aes(ymin = exp_mean - 2*exp_sd, ymax = exp_mean + 2*exp_sd), position = position_dodge(width = 0.2)) + 
  scale_x_discrete(limits = c('PBS_PYO','PYO_PBS','PCA_PYO','PCN_PYO'), labels = parse(text = labels )) + 
  scale_color_manual(breaks = c(F,T), labels = c('-DNA','+DNA'), values = colorblind_palette)+
  labs(x = NULL, y = 'Absorbance at 690nm', color = NULL) + 
  theme(axis.text.x = element_text( size=10,angle = 45, hjust = 1))

plot_ctDNA_690
```

Beautiful.

## Fluorescence

Since we used fuorescence data to try to look for binding of reduced phenazines to ctDNA, let's also see what that data look like:

```{r}
df_fluor <- read_csv("data/2019_10_18_ET_ctDNA_fluor_360ex_460em_1.csv") %>%
    gather(key = 'well', value = 'intensity', -min)

df_fluor_meta <- left_join(df_fluor, metadata, by = 'well')

ggplot(df_fluor_meta, aes(x = min, y = intensity, color = DNA)) + geom_point() + facet_wrap(~red_ox)
```

Interesting, sort of looks like no DNA yields higher signal...let's aggregate and look at pointranges:

```{r}
df_fluor_well <- df_fluor_meta %>% group_by(red_ox, rep, DNA) %>% summarise(well_mean = mean(intensity), well_sd = sd(intensity))

ggplot(df_fluor_well, aes(x = red_ox, y = well_mean, color = DNA)) + geom_pointrange(aes(ymin = well_mean - 2*well_sd, ymax = well_mean + 2*well_sd), position = position_jitter(width = 0.1))
```

Here we can definitely see that the values are lower + DNA. However, they are lower across the board...even for oxidized PYO, which has no fluorescence. Therefore I think this is mostly explained by the observed bias in the platereader, which produces lower fluorescence values in this for columns 4-6 than 1-3...

# Conclusions

There is no evidence that DNA binding inhibits interphenazine ET to PYOox. Therefore inter phenazine ET may be compatible with DNA binding and retention of PYOox, while still facilitating rapid ET to O2.

## Future directions?

One other experiment that may be worth trying is to test the reaction rates of PCAred and PCNred in the presence or absence of PYOox (and perhaps +/- ctDNA). My prediction is that the rate will be much faster in the presence of PYO, since interphenazine ET is very fast. This is something that I have thought about bringing up in the discussion of the paper, since it might sort of make sense that the system efficiently transfers electrons to oxygen, while allowing PYO to be retained and PCN/PCAred to be reoxidized quickly...may be too subtle of a point for the reader, but I think it's pretty interesting and straightforward.

```{r}

sessionInfo()

```
