---
title: "Agarose ctDNA"
subtitle: "Analysis"
author: "Scott Saunders"
fontsize: 12pt
date: "08_13_19 "
output:
  html_document:
    self_contained: false
    toc: true
    code_folding: hide
  github_document:
    pandoc_args: --webtex
---

```{r setup, echo=T, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom) 
library(modelr) 
library(viridis)
library(lubridate)
library(hms)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center")

source("../../tools/echem_processing_tools.R")
source("../../tools/plotting_tools.R")

theme_set(theme_1())
```

# Data
```{r}
# Agarose
gc_agarose <- read_csv("../../06_06_19_agarose_PYO_2/Processing/06_06_19_processed_gc_max_agarose.csv")
swv_agarose <- read_csv("../../06_06_19_agarose_PYO_2/Processing/06_06_19_processed_swv_max_agarose.csv")
swvGC_agarose <- read_csv("../../06_06_19_agarose_PYO_2/Processing/06_06_19_processed_swvGC_agarose.csv")

# Agarose + ctDNA

gc_ctDNA <- read_csv("../processing/08_13_19_processed_gc_max_ctDNA.csv")
swv_ctDNA <- read_csv("../processing/08_13_19_processed_swv_max_ctDNA.csv")
swvGC_ctDNA <- read_csv("../processing/08_13_19_processed_swvGC_ctDNA.csv")
```

# GC time course

## Overview
```{r}

df_gc <- bind_rows(
  gc_agarose %>% mutate(condition = 'agarose'), 
  gc_ctDNA %>% mutate(condition = 'ctDNA')
  ) %>% 
  group_by(condition, reactor) %>% 
  mutate(min_time = min(minutes)) %>% 
  mutate(norm_time = minutes - min_time)

ggplot(df_gc, aes(x = norm_time, y = max_current, color = reactor, shape = electrode)) + geom_point() + facet_wrap(~condition)
```

## Transfer runs
```{r}
ggplot(df_gc %>% filter(norm_time <100) %>% filter(reactor != 'soak' & reactor !='soak_2'), aes(x = norm_time, y = max_current, color = condition, shape = electrode)) + 
  geom_point()+
  facet_wrap(~electrode)

```
Looks like their may actually be a small difference especially when normalizing by the starting values...

## Soak Runs
```{r}
ggplot(df_gc %>% filter(norm_time <100) %>% filter(reactor == 'soak' | reactor =='soak_2'), aes(x = norm_time, y = max_current, color = condition, shape = electrode)) + 
  geom_point()+
  facet_wrap(~electrode)

```
Hard to tell without normalization / letting it go longer...but ctDNA may actually be slower. 

# SWV time course

## Overview
```{r}

df_swv <- bind_rows(
  swv_agarose %>% mutate(condition = 'agarose'), 
  swv_ctDNA %>% mutate(condition = 'ctDNA')
  ) %>% 
  group_by(condition, reactor) %>% 
  mutate(min_time = min(minutes)) %>% 
  mutate(norm_time = minutes - min_time)

ggplot(df_swv, aes(x = norm_time, y = max_current, color = reactor, shape = electrode)) + geom_point() + facet_wrap(~condition)
```

## Transfer Runs

```{r}
ggplot(df_swv %>% filter(reactor != 'soak' & reactor !='soak_2') %>% filter(norm_time<100), aes(x = norm_time, y = max_current, color = reactor, shape = electrode)) + geom_point() + facet_wrap(~electrode)
```
Again, looks like the ctDNA may actually be slower, especially after normalizing. Transfer 2 / 3 are replicates and look pretty similar, so that's good. 

## Soak Runs

```{r}
ggplot(df_swv %>% filter(reactor == 'soak' | reactor =='soak_2'), aes(x = norm_time, y = max_current, color = reactor, shape = electrode)) + geom_point() + facet_wrap(~electrode)
```

Same as before...a little hard to tell without normalization or running longer. ctDNA is possibly slower. 

# SWV vs. GC

## Overview

```{r}
df_swv_gc <- bind_rows(
  swvGC_agarose %>% mutate(condition = 'agarose'), 
  swvGC_ctDNA %>% mutate(condition = 'ctDNA')
  )

ggplot(df_swv_gc, aes(x = max_current_swv, y = max_current_gc, color = reactor)) + geom_point() + facet_wrap(~electrode_swv) + scale_color_viridis_d()+
  geom_smooth(data = df_swv_gc %>% filter(rep <= 10), method = 'lm')
```

Overall slopes look pretty similar. Transfer_2 is an obvious outlier with a shallower slope. 

# Conclusions

* Need to actually perform Dphys and Dap calculations. 
* Probably need to repeat both experiments (and maybe glycerol condition too)...acquisitions take a while though. How long to run??
